---
title: "Secondary Processing"
output: html_document
date: "2024-03-19"
---

**Important note**: The processing of subsets are performed in separate Rmd files. However, the processing of all subsets are gathered in the same Rmd for the final version (V2) of whole dataset.There is a version 3, which is created between V2p3 and V2p4, containing only hashtag single positive cell. Version 3 dataset was not presented.

In this Rmd, secondary cleaning will be performed. The preliminary annotation can be performed along with cleaning or after cleaning. 
I have separated annotation from secondary processing for pacing purposes. The codes in this Rmd recapitulates steps performed in processing V2 (i.e creation of V2p3 dataset from prediction filtered whole dataset.)

```{r}
library(Seurat)
library(tidyverse)
library(harmony)
library(DoubletFinder)
library(miQC)
library(SoupX)
library(data.table)
library(SeuratWrappers) #For miQC
```

```{r}
indir = "./R_outs/"
outdir = "./R_outs/"

```

# Part 1 process the subsets

## 1.1 process lymphoid


```{r}
load(paste0(outdir,"lymphoid.Robj"))
```

### 1.1.1 Iteration 1 

```{r}
lymphoid <- NormalizeData(lymphoid,assay = "RNA", normalization.method = "LogNormalize", scale.factor = 10000)
lymphoid <- FindVariableFeatures(object = lymphoid, selection.method = 'mean.var.plot', mean.cutoff = c(0.025, 5), dispersion.cutoff = c(0.55, Inf))
VariableFeatures(lymphoid) <-  VariableFeatures(lymphoid) [!grepl("^Tra|^Trab|^Igh|^Igk|Mamu-a", VariableFeatures(lymphoid))]
VariableFeaturePlot(lymphoid,selection.method = "mean.var.plot")

lymphoid <- ScaleData(object = lymphoid, features = VariableFeatures(object = lymphoid), vars.to.regress = c("nCount_RNA", "percent.mt"))
lymphoid <- RunPCA(object = lymphoid,
                features =  VariableFeatures(object = lymphoid),
                dims = 1:50)
gc()

ElbowPlot(lymphoid,ndims = 50)
```


```{r}
lymphoid <- RunHarmony(object = lymphoid, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
lymphoid <- RunUMAP(lymphoid,dims = 1:20,reduction = "harmony")
lymphoid <- RunTSNE(lymphoid,dims = 1:20,reduction = "harmony")

lymphoid <- FindNeighbors(lymphoid, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  lymphoid <- FindClusters(object = lymphoid, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r}
Idents(lymphoid) <- lymphoid$RNA_snn_res.0.1

DimPlot(lymphoid,label=T)

FeaturePlot(lymphoid,features = "Ncr1",label=T)
FeaturePlot(lymphoid,features = "Cd8a",label=T)
FeaturePlot(lymphoid,features = "Cd4",label=T)
FeaturePlot(lymphoid,features = "Cd3e",label=T)
FeaturePlot(lymphoid,features = "Adgre1",label=T)
```

```{r remove doublet-like clusters}
lymphoid <- subset(lymphoid,idents = c(0,2,3))
```

```{r}
save(lymphoid,file=paste0(outdir,"lymphoid_V2.Robj")) # make sure you keep the doublet-like removed version as a different version
```

### 1.1.2 Iteraction 2

```{r}
lymphoid <- FindVariableFeatures(object = lymphoid, selection.method = 'mean.var.plot', mean.cutoff = c(0.025, 5), dispersion.cutoff = c(0.55, Inf))
VariableFeatures(lymphoid) <-  VariableFeatures(lymphoid) [!grepl("^Tra|^Trab|^Igh|^Igk|Mamu-a", VariableFeatures(lymphoid))]
VariableFeaturePlot(lymphoid,selection.method = "mean.var.plot")

lymphoid <- ScaleData(object = lymphoid, features = VariableFeatures(object = lymphoid), vars.to.regress = c("nCount_RNA", "percent.mt"))
lymphoid <- RunPCA(object = lymphoid,
                features =  VariableFeatures(object = lymphoid),
                dims = 1:50)
gc()
lymphoid <- RunHarmony(object = lymphoid, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
lymphoid <- RunUMAP(lymphoid,dims = 1:20,reduction = "harmony")
lymphoid <- RunTSNE(lymphoid,dims = 1:20,reduction = "harmony")

lymphoid <- FindNeighbors(lymphoid, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  lymphoid <- FindClusters(object = lymphoid, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r}
Idents(lymphoid) <- lymphoid$RNA_snn_res.1
DimPlot(lymphoid,label=T)
```

```{r}
DotPlot(lymphoid, features = c("Adgre1","Ncr1","Tyrobp","Xcl1","Cd3e","Cd8a","Cd4","Mki67"))+RotatedAxis()
# remove cluster 14

FeaturePlot(lymphoid,features = "Ncr1",label=T)
FeaturePlot(lymphoid,features = "Cd8a",label=T)
FeaturePlot(lymphoid,features = "Cd4",label=T)
FeaturePlot(lymphoid,features = "Cd3e",label=T)
FeaturePlot(lymphoid,features = "Adgre1",label=T)
FeaturePlot(lymphoid,features = "Mki67",label=T)
```

```{r}
lymphoid <- subset(lymphoid,idents = c(0:13,15:16)) 
save(lymphoid,file=paste0(outdir,"lymphoid_V2.Robj"))
```

### 1.1.3 Iteration 3

```{r}
lymphoid <- FindVariableFeatures(object = lymphoid, selection.method = 'mean.var.plot', mean.cutoff = c(0.025, 5), dispersion.cutoff = c(0.55, Inf))
VariableFeatures(lymphoid) <-  VariableFeatures(lymphoid) [!grepl("^Tra|^Trab|^Igh|^Igk|Mamu-a", VariableFeatures(lymphoid))]
VariableFeaturePlot(lymphoid,selection.method = "mean.var.plot")

lymphoid <- ScaleData(object = lymphoid, features = VariableFeatures(object = lymphoid), vars.to.regress = c("nCount_RNA", "percent.mt"))
lymphoid <- RunPCA(object = lymphoid,
                features =  VariableFeatures(object = lymphoid),
                dims = 1:50)
gc()
lymphoid <- RunHarmony(object = lymphoid, group.by.vars = c("orig.ident"), max.iter.harmony = 20)

lymphoid <- RunUMAP(lymphoid,dims = 1:15,reduction = "harmony")
lymphoid <- RunTSNE(lymphoid,dims = 1:15,reduction = "harmony")

lymphoid <- FindNeighbors(lymphoid, dims = 1:15,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  lymphoid <- FindClusters(object = lymphoid, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()

lymphoid <- JoinLayers(lymphoid)
save(lymphoid,file=paste0(outdir,"lymphoid_V2.Robj"))
```


## 1.2 process myeloids


```{r}
load(paste0(outdir,"myeloid.Robj"))
```

### 1.2.1 Iteraction 1


```{r}
myeloid <- NormalizeData(myeloid,assay = "RNA", normalization.method = "LogNormalize", scale.factor = 10000)
myeloid <- FindVariableFeatures(object = myeloid, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(myeloid) <-  VariableFeatures(myeloid) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(myeloid))]
VariableFeaturePlot(myeloid)

myeloid <- ScaleData(object = myeloid, features = VariableFeatures(object = myeloid), vars.to.regress = c("nCount_RNA", "percent.mt"))
myeloid <- RunPCA(object = myeloid,
                features =  VariableFeatures(object = myeloid),
                dims = 1:50)
gc()

ElbowPlot(myeloid ,ndims = 50)
```

```{r}
myeloid <- RunHarmony(object = myeloid , group.by.vars = c("orig.ident"), max.iter.harmony = 20)
myeloid <- RunUMAP(myeloid ,dims = 1:40,reduction = "harmony")
myeloid <- RunTSNE(myeloid ,dims = 1:40,reduction = "harmony")

myeloid <- FindNeighbors(myeloid , dims = 1:40,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  myeloid  <- FindClusters(object = myeloid , resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
myeloid <- JoinLayers(myeloid)
```

```{r}
Idents(myeloid) <- myeloid$RNA_snn_res.0.1

DimPlot(myeloid,label = T)
DimPlot(myeloid, reduction = "umap",group.by = "orig.ident",shuffle = T)
DimPlot(myeloid, reduction = "umap")
FeaturePlot(myeloid,features = "Cd3e",reduction = "umap")
FeaturePlot(myeloid,features = "Ncr1",reduction = "umap")
FeaturePlot(myeloid,features = "Adgre1",reduction = "umap")
FeaturePlot(myeloid,features = "Mki67",reduction = "umap")
FeaturePlot(myeloid,features = "S100a8",reduction = "umap")
FeaturePlot(myeloid,features = "Nos2",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Adgre1",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Ccr2",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Ccl2",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Cx3cr1",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Il1r2",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Itga4",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
FeaturePlot(myeloid,features = "Mertk",reduction = "tsne",label = T)+scale_colour_gradientn(colors = c("grey", "orange", "red"))
```

```{r}
DotPlot(myeloid,features = c("Cd3e","Cd8a","Cd4","Adgre1","Itgam","Mki67","Cd200r3","Nos2","Cx3cr1","Xcr1","Mertk","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"))+RotatedAxis() #Remove cluster 6 as a doublet cluster
```

```{r}
myeloid <- subset(myeloid,idents = c(0:5))
save(myeloid,file=paste0(outdir,"myeloid_V2.Robj"))
```

### 1.2.2 Iteraction 2

```{r}
myeloid <- NormalizeData(myeloid,assay = "RNA", normalization.method = "LogNormalize", scale.factor = 10000)
myeloid <- FindVariableFeatures(object = myeloid, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(myeloid) <-  VariableFeatures(myeloid) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(myeloid))]
VariableFeaturePlot(myeloid)

myeloid <- ScaleData(object = myeloid, features = VariableFeatures(object = myeloid), vars.to.regress = c("nCount_RNA", "percent.mt"))
myeloid <- RunPCA(object = myeloid,
                features =  VariableFeatures(object = myeloid),
                dims = 1:50)
gc()
myeloid <- RunHarmony(object = myeloid , group.by.vars = c("orig.ident"), max.iter.harmony = 20)
myeloid <- RunUMAP(myeloid ,dims = 1:40,reduction = "harmony")
myeloid <- RunTSNE(myeloid ,dims = 1:40,reduction = "harmony")

myeloid <- FindNeighbors(myeloid , dims = 1:40,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  myeloid  <- FindClusters(object = myeloid , resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r}
DotPlot(myeloid,features = c("Cd3e","Cd8a","Cd4","Adgre1","Itgam","Mki67","Cd200r3","Nos2","Cx3cr1","Xcr1","Mertk","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"))+RotatedAxis()

DimPlot(myeloid,label=T) # remove 18
```

```{r}
myeloid <- subset(myeloid,idents = c(0:17,19:22))
save(myeloid,file=paste0(outdir,"myeloid_V2.Robj"))
```

### 1.2.3 Iteraction 3

```{r}
myeloid <- NormalizeData(myeloid,assay = "RNA", normalization.method = "LogNormalize", scale.factor = 10000)
myeloid <- FindVariableFeatures(object = myeloid, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(myeloid) <-  VariableFeatures(myeloid) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(myeloid))]
VariableFeaturePlot(myeloid)

myeloid <- ScaleData(object = myeloid, features = VariableFeatures(object = myeloid), vars.to.regress = c("nCount_RNA", "percent.mt"))
myeloid <- RunPCA(object = myeloid,
                features =  VariableFeatures(object = myeloid),
                dims = 1:50)
gc()
ElbowPlot(myeloid)
```

```{r}
myeloid <- RunHarmony(object = myeloid , group.by.vars = c("orig.ident"), max.iter.harmony = 20)
myeloid <- RunUMAP(myeloid ,dims = 1:20,reduction = "harmony")
myeloid <- RunTSNE(myeloid ,dims = 1:20,reduction = "harmony")

myeloid <- FindNeighbors(myeloid , dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  myeloid  <- FindClusters(object = myeloid , resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r}
DimPlot(myeloid,label=T)
DotPlot(myeloid,features = c("Cd3e","Cd8a","Cd4","Adgre1","Itgam","Mki67","Cd200r3","Nos2","Cx3cr1","Xcr1","Mertk","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"))+RotatedAxis()
```

```{r}
myeloid <- JoinLayers(myeloid)
save(myeloid,file=paste0(outdir,"myeloid_V2.Robj"))
```

## 1.3 process stromal

```{r}
load(paste0(outdir,"stromal.Robj"))
```


### 1.3.1 Iteration 1

```{r}
stromal <- NormalizeData(stromal,assay = "RNA", normalization.method = "LogNormalize", scale.factor = 10000)
stromal <- FindVariableFeatures(object = stromal, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(stromal) <-  VariableFeatures(stromal) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(stromal))]
VariableFeaturePlot(stromal)

stromal <- ScaleData(object = stromal, features = VariableFeatures(object = stromal), vars.to.regress = c("nCount_RNA", "percent.mt"))
stromal <- RunPCA(object = stromal,
                features =  VariableFeatures(object = stromal),
                dims = 1:50)
gc()
ElbowPlot(stromal)
```

```{r}
stromal <- RunHarmony(object = stromal , group.by.vars = c("orig.ident"), max.iter.harmony = 20)
stromal <- RunUMAP(stromal ,dims = 1:15,reduction = "harmony")
stromal <- RunTSNE(stromal ,dims = 1:15,reduction = "harmony")

stromal <- FindNeighbors(stromal, dims = 1:15,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  stromal  <- FindClusters(object = stromal , resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
DimPlot(stromal,label=T)
stromal <- JoinLayers(stromal)
save(stromal,file=paste0(outdir,"stromal_V2.Robj"))
```

```{r}
DotPlot(stromal,features=c("Cd3e","Cd8a","Cd4","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3","Col5a2","Mki67","Myct1","Egfr","Pecam1","Ptprc"))+RotatedAxis()
```

```{r}
stromal <- subset(stromal,idents = c(0:4,6:11,13))
```

### 1.3.2 Iteration 2


```{r}
stromal <- FindVariableFeatures(object = stromal, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(stromal) <-  VariableFeatures(stromal) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(stromal))]
VariableFeaturePlot(stromal)

stromal <- ScaleData(object = stromal, features = VariableFeatures(object = stromal), vars.to.regress = c("nCount_RNA", "percent.mt"))
stromal <- RunPCA(object = stromal,
                features =  VariableFeatures(object = stromal),
                dims = 1:50)
gc()
ElbowPlot(stromal)
```

```{r}
stromal <- RunHarmony(object = stromal , group.by.vars = c("orig.ident"), max.iter.harmony = 20)
stromal <- RunUMAP(stromal ,dims = 1:5,reduction = "harmony")
stromal <- RunTSNE(stromal ,dims = 1:5,reduction = "harmony")

stromal <- FindNeighbors(stromal, dims = 1:5,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  stromal  <- FindClusters(object = stromal , resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
DimPlot(stromal,label=T)
stromal <- JoinLayers(stromal)
```

```{r}
DotPlot(stromal,features=c("Cd3e","Cd44","Ncam1","Col5a2","Mki67","Myct1","Egfr","Pecam1","Ptprc","Il6","Fap","S100a9", "S100a8", "G0s2", "Nampt", "Rgs2", "Sod2","Fabp7", "Vegfa", "Spp1","Rgs5", "Igfbp7", "Sparcl1", "Mcam", "Myh11","A2m", "Cpe", "Adamts1", "Acta2"))+RotatedAxis()
```

```{r}
save(stromal,file=paste0(outdir,"stromal_V2.Robj"))
```

# Part 2 Merge subsets back to whole dataset (Optional)

Ideally, annotation should be performed at the subset level. Thus, it can be less efficient to merge back the whole dataset before completing preliminary annotation, of which aim is not to create the finalized annotation but to generate a placeholder for the annotations.

The codes are pasted for documentation purposes.

```{r}
whole <- merge(x = myeloid, y = c(lymphoid,stromal))
```

```{r}
head(whole@meta.data)
```

```{r}
whole <- NormalizeData(object = whole, normalization.method = "LogNormalize", scale.factor = 10000)
#there is no need to normalize again; normalization is cell specific
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(whole))]
VariableFeaturePlot(whole)

whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA", "percent.mt"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```

```{r}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
whole <- RunUMAP(whole,dims = 1:40,reduction = "harmony")
whole <- RunTSNE(whole,dims = 1:40,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:40,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
DimPlot(whole,label=T)
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Adgre1","Lyz2","Xcr1","Ncr1","Siglech","Mki67","S100a8","Tpsb22","Fcer1a","Cd200r3","Fcer2a","Flt3","Pecam1","Myct1"))+RotatedAxis()
```
```{r}
whole <- JoinLayers(whole)
```


```{r}
FeaturePlot(whole,features = c("Cd3e"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Ncr1"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Adgre1"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Siglech"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Flt3"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Xcr1"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Mki67"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("S100a8"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Fcer1a"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
FeaturePlot(whole,features = c("Myct1"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
```

```{r}
FeaturePlot(whole,features = c("Cd200r3"),label=T)+scale_colour_gradientn(colors = c("grey", "red", "purple"))
```

```{r}
save(whole,file = paste0(outdir,"whole_V2p3.Robj"))
```



