---
title: "GSEA"
output: html_document
date: "2024-03-19"
---

**Note**: the original file is Functional_Analysis_and_Visualization.Rmd, which contains the full codes for completing the analysis for all compartment of interest. This Rmd is for demonstration purpose.The GO analysis section is removed in this version.

Note that I might have taken advantadge of the global environment and used the markers that were generated in another Rmd file that was opened simultanouesly as Functional_Analysis_and_Visualization.Rmd. Thus, the original Rmd may have the loading steps missing.

```{r}
library(msigdbr) # for retreiving geneset
library(fgsea)
library(org.Mm.eg.db) # in case gene name conversion is needed.
library(Seurat) # you know why
library(tidyverse) # base package for data organization and plotting
library(RColorBrewer) # for making palettes
library(gridExtra) # for arranging multiple plots
library(data.table) # well...
```

```{r}
indir <- "./" # These are not the actual indir and outdir
outdir <- "./"
```

# Step 0 prepare helper functions

```{r fgsea_res}
fgsea_res <- function(cell_type_vector, DE_genes, choose_n = 100, choose_tops = F,gene_set = fgsea_sets_Hallmark,mode = "combined"){
  fgsea_res <- list()
  DE_selected <- DE_genes %>% mutate(metric = sign(avg_log2FC)*(-log10(p_val_adj))) %>% filter(!is.infinite(metric))
  if (mode == "simple"){
    DE_selected <- DE_genes %>% mutate(metric = avg_log2FC) %>% filter(!is.infinite(metric))
  }
  for (i in cell_type_vector) {
    DE <- DE_selected[DE_selected$cluster == i,] %>% arrange(desc(metric)) %>% select(gene,metric)
    if (choose_tops){
      DE_top <- DE %>% group_by(cluster) %>% slice_max(order_by = metric, n = choose_n)
      DE_tail <- DE %>% group_by(cluster) %>% slice_min(order_by = metric, n = choose_n)
      DE <- rbind(DE_top,DE_tail)
    }
    DE <- DE %>% arrange(desc(metric)) %>% select(gene,metric)
    rankings <- DE$metric
    names(rankings) <- DE$gene
    #print(rankings)
    plot(rankings)
    fgseaRes <- fgsea(gene_set, stats = rankings,nPermSimple = 10000)
    fgseaResTidy <- fgseaRes %>%
      as_tibble() %>% filter(padj < 0.05 ) %>%
      arrange(desc(NES))
  fgsea_res[[i]] <- fgseaResTidy
  fwrite(fgseaResTidy, file=paste0(i,"_fgsea.tsv"), sep="\t", sep2=c("", " ", ""))
  }
   return(fgsea_res)
}
```

```{r helper function for generate palette that matches seurat color coding}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```


# Step 1 prepare gene sets

You can also load gene sets from Rda, but I loaded those from `msigdbr` package.


```{r prepare gene sets}
H_df<- msigdbr(species = "Mus musculus", category = "H")
Imm_df <- msigdbr(species = "Mus musculus", category = "C7",subcategory = "IMMUNESIGDB")
Celltype_df <- msigdbr(species = "Mus musculus", category = "C8")
C2_KEGG <- msigdbr(species = "Mus musculus", category = "C2",subcategory = "CP:KEGG")
C2_React <- msigdbr(species = "Mus musculus", category = "C2",subcategory = "CP:REACTOME")

fgsea_sets_Hallmark <- H_df %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_Imm <- Imm_df %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_KEGG <- C2_KEGG %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_REACT <- C2_React %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_Celltype <- Celltype_df %>% split(x = .$gene_symbol, f = .$gs_name)
```

# Step 2 run fgsea analysis over intended

```{r}
NK_marker <- read.table(file = paste0(indir,"NK_annotation_markers.tsv"), sep = "\t",header = T,row.names = F)
```


```{r}
NK_fgsea_hallmark <- fgsea_res(as.character(unique(NK_markers$cluster)),NK_markers,mode = "simple")

NK_fgsea_react <- fgsea_res(as.character(unique(NK_markers$cluster)),NK_markers,gene_set = fgsea_sets_REACT,mode = "simple")

NK_fgsea_KEGG <- fgsea_res(as.character(unique(NK_markers$cluster)),NK_markers,gene_set = fgsea_sets_KEGG,mode = "simple")

NK_fgsea_Imm <- fgsea_res(as.character(unique(NK_markers$cluster)),NK_markers,gene_set = fgsea_sets_Imm,mode = "simple")

NK_fgsea_CT <- fgsea_res(as.character(unique(NK_markers$cluster)),NK_markers,gene_set = fgsea_sets_Celltype,mode = "simple")
```

# Step 3 consolidate GSEA results from all cell types into a table

## NK hallmarks


```{r}
Hallmark_vector <- names(fgsea_sets_Hallmark)
celltype_vector <- as.character(unique(NK_markers$cluster))
table_res_NK <- data.frame(pathway = Hallmark_vector)
for (i in celltype_vector){
  table_tmp <- NK_fgsea_hallmark[[i]]
  table_celltype <- table_tmp[table_tmp$pathway %in% Hallmark_vector,c("pathway","NES")]
  colnames(table_celltype) <- c("pathway",i)
  table_res_NK <- full_join(table_res_NK,table_celltype,by = "pathway")
}
```

```{r}
table_res_NK[is.na(table_res_NK)] = 0
table_res_NK <- table_res_NK[apply(table_res_NK[,2:6],1,var) != 0,]
```

```{r}
write.table(table_res_NK,file = paste0(outdir,"fgsea_NK_Hallmark_consolidated_filtered.tsv"),sep = "\t",row.names = F,quote = F)
```

# Step 4 visualization of fgsea results as a heatmap

```{r}
table_res <- read.table(file = paste0(outdir,"fgsea_NK_Hallmark_consolidated_filtered.tsv"),sep = "\t",header = T,row.names = F)
```

```{r}
n_celltype = (ncol(table_res) - 1)
ncols = ncol(table_res)
pal <- gg_color_hue(n_celltype) # 
names(pal) <- colnames(table_res)[2:ncols]
# match the order in seurat dimplot to make color-coding consistent
col_gsea = list(
  Cell_type = pal
)
ha <- HeatmapAnnotation(
   Cell_type = colnames(table_res)[2:ncols],
  col = col_gsea
)
gsea_mat <- as.matrix(table_res[,2:ncols])
rownames(gsea_mat) <- table_res$pathway
Heatmap(gsea_mat,name = "GSEA",top_annotation = ha,cluster_rows = F,cluster_columns = T,show_row_names = TRUE,row_names_gp = grid::gpar(fontsize = 5),
        show_column_names = F,row_names_side = "left",col = colorRamp2(c(-2,0,2),c("darkblue","beige","red")))
```

```{r}
pdf(paste0(outdir,"GSEA_HALLMARK_heatmap.pdf")) # change file name properly
Heatmap(gsea_mat,name = "GSEA",top_annotation = ha,cluster_rows = F,cluster_columns = T,show_row_names = TRUE,row_names_gp = grid::gpar(fontsize = 5),
        show_column_names = F,row_names_side = "left",col = colorRamp2(c(-2,0,2),c("darkblue","beige","red")))
dev.off()
```

