---
title: "Within-cluster DE"
output: html_document
date: "2024-03-19"
---

```{r}
library(Seurat)
library(tidyverse)
library(data.table)
library(MAST) # for DE
library(EnhancedVolcano) # for Volcano plot
```

**Note**: Cross-condition DE is originally performed using R scripts. This Rmd shows one example.

# Part 1: Within-cell type DE

**Note**: This part is dependent on settlement of cell type definition and preliminary annotation.

## 1.1 load whole object

For the simplicity of object loading, only the whole object is loaded. Subset will be created using the cell type label.

```{r}
indir = "./R_outs/"
outdir = "./R_outs/"
```

```{r}
load(paste0(outdir,"whole_V2p4.Robj"))
```

## 1.2 DE of subsets

### 1.2.1 perform DE among CAF

```{r}
CAF <- subset(whole,subset = celltype == "CAF")
```

```{r}
table(CAF$annotation)
```


```{r}
Idents(CAF) <- CAF$annotation
CAF_markers <- FindAllMarkers(CAF)
```

```{r}
write.table(CAF_markers,
              paste0(outdir,"CAF_annotation_markers.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```


### 1.2.2 DE among DC

```{r}
DC <- subset(whole,subset = celltype == "DC")
table(DC$annotation)
```

```{r}
Idents(DC) <- DC$annotation

markers <- FindAllMarkers(DC)

write.table(markers,
              paste0(outdir,"DC_annotation_markers.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```

### 1.2.3 DE Neutrophils

```{r}
Neutrophil <- subset(whole,subset = celltype == "Neutrophil")
table(Neutrophil$annotation)
```

```{r}
Idents(Neutrophil) <- Neutrophil$annotation

markers <- FindAllMarkers(Neutrophil)

write.table(markers,
              paste0(outdir,"Neutrophil_annotation_markers.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```

### 1.2.4 DE among mono/mac

```{r}
mono_mac <- subset(whole,subset = celltype == "Mono/Mac")

```

```{r}
table(mono_mac$annotation_V2)
```


```{r}
Idents(mono_mac) <- mono_mac$annotation_V2

markers <- FindAllMarkers(mono_mac)

write.table(markers,
              paste0(outdir,"mono_mac_annotation_markers_V2.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```

### 1.2.5 DE of lymphoid cell types


```{r}
NK <- subset(whole, subset = celltype == "NK")
T_cell <- subset(whole, subset = celltype == "T cell")
```

```{r}
Idents(NK) <- NK$annotation
NK_markers <- FindAllMarkers(NK)
write.table(NK_markers,file = paste0(outdir,"NK_annotation_markers.tsv"),sep = "\t",row.names = F)
```


```{r}
Idents(T_cell) <- T_cell$annotation
T_cell_markers <- FindAllMarkers(T_cell)
write.table(T_cell_markers,file = file = paste0(outdir,"T_cell_annotation_markers.tsv"),sep = "\t",row.names = F)
```


# Part 2: Cross-condition DE for a list of clusters within an object

**Note**: the name of the object is hard-coded, so whenever the object changes, the object name needs to be updated.

Be aware of the annotation/identity classes that you are iterating the DE over.

## 2.1 define function

```{r}
de_clusters <- function(current_cluster) {
  groups <- paste0(conditions,"_",current_cluster)
  both_markers <- FindMarkers(neutrophil, ident.1=groups[2], ident.2=groups[1], test.use="MAST", min.pct=0.00, logfc.threshold = 0) # change name of object 
  marker_genes <- row.names(both_markers)
  mutate(both_markers, gene = marker_genes, cluster = current_cluster)

}
```

## 2.2 prepare object

```{r}
neutrophil$group <- gsub(" ","_",paste0(neutrophil$orig.ident,"_",neutrophil$celltype))
conditions <- unique(neutrophil$orig.ident)
clusters <- "Neutrophil"
plts <- list()
padj_thres = 10e-2
log2FC_thres = 1
Idents(neutrophil) <- neutrophil$group
```

## 2.3 run DE over a list of clusters

```{r}
for (i in clusters) {
  output_file <- paste0(outdir,paste0('ZBTB46_vs_Control_', i, '.tsv'))
  output_file_flt <- paste0(outdir,paste0('ZBTB46_vs_Control_', i, '_filtered.tsv'))
  de_results_tbl <- lapply(i, de_clusters) 
  de_results_tbl <- de_results_tbl %>% bind_rows() %>%
			arrange(desc(abs(avg_log2FC))) %>%
			mutate(threshold = case_when(
			(p_val_adj < padj_thres) & (abs(avg_log2FC) > log2FC_thres) ~ "Pass",
			.default = "Fail"
			))
  write.table(de_results_tbl, output_file,sep="\t",row.names=T)
  
  de_results_tbl <- de_results_tbl %>% 
			filter(p_val_adj < padj_thres) %>%
			filter(abs(avg_log2FC) > log2FC_thres)
			
  write.table(de_results_tbl, output_file_flt,sep="\t",row.names=T)
}
 
```

# Part 3 summary and visualization of cross-condition DE

Since across cluster DE is performed one v.s multiple, the summary and visualization is not as straightforward as those for cross-condition DE.

These are just demonstrations.

## 3.1 count number of DE in each cluster

```{r}
indir <- "R_outs"
outdir <- "R_outs"
```

```{r}
lym.DE.files <- list.files(path = "R_outs",pattern = "ZBTB46_vs_Control_lymphoid")
i = 1
clust_name <- gsub(".tsv","",gsub("ZBTB46_vs_Control_lymphoid","",lym.DE.files[i*2]))
```

```{r define helper function}
calculate_n_DE <- function(file_list,celltype){
  n_clust <- (length(file_list))/2
  # ideally we want to do apply, but I'm dumb, so I'll do it in a less elegant way
  cluster <- character(n_clust)
  total <- integer(n_clust)
  filtered <- integer(n_clust)
  for (i in seq(1:n_clust)){
    clust_name[i] <- gsub("_|.tsv","",gsub(paste0("ZBTB46_vs_Control_",celltype),"",file_list[i*2]))
    #clust name automatically detected. unfiltered ones are expected to be the even-numbered item of the vector
    total[i] <- nrow(read.table(paste0(indir,"/",file_list[i*2]),sep = "\t" ))
    filtered[i] <- nrow(read.table(paste0(indir,"/",file_list[i*2-1]),sep = "\t" ))

    
  }
  return(n_DE_DF <- data.frame(cluster = clust_name,n_total = total,n_pass = filtered))
}
```

```{r generate results}
n_DE_lym <- calculate_n_DE(lym.DE.files,"lymphoid")

ggplot(n_DE_lym,aes(x=reorder(cluster, as.integer(cluster)),y=n_pass))+
  geom_bar(stat = "identity", position = 'dodge')+xlab("cluster number")+ylab("number of DE genes with padj < 0.01")+ggtitle("lymphoid at resolution 0.6")
```


## 3.2 make volcano plot

```{r load cross-condition DE result}
# You'll need to load all results
DE_df <- read.table(paste0(outdir,"ZBTB46_vs_Control_stromal_0.tsv"),header = T)
```

```{r enhanced volcano plot}
EnhancedVolcano(DE_df,
    lab = rownames(DE_df),
    x = 'avg_log2FC',
    y = 'p_val_adj',
    pCutoff = 1e-2,
    ylab = "adjusted P-value")
g <- EnhancedVolcano(DE_df,
    lab = rownames(DE_df),
    x = 'avg_log2FC',
    y = 'p_val_adj',
    pCutoff = 1e-2,
    ylab = "adjusted P-value",
    title = "Treatment v.s. Control")

ggsave(filename =  "Stromal_volcano_res02_c0.pdf",plot = g, units = "cm",height = 20,width = 15)
```





