---
title: "T cells analysis"
output: html_document
---

```{r}
library(Seurat)
library(tidyverse)
library(harmony)
```


```{r}
indir = "./R_outs/" #Can be made into a command-line input
outdir = "./R_outs/"
```

```{r}
load(paste0(outdir,"whole_processed_wHarmony.Robj"))
```

## Divide two subsets: CD3e high and Cd3e low for further clustering


```{r}
VlnPlot(whole, features = "Cd3e",group.by="RNA_snn_res.0.1")
```

```{r}
Idents(whole) <- whole$RNA_snn_res.0.1
CD3E_high <- subset(whole, idents = c(0,1,2,4,8))
CD3E_low <- subset(whole, idents = c(3,5,6,7,9))
```

```{r}
table(CD3E_high$RNA_snn_res.0.1)
table(CD3E_low$RNA_snn_res.0.1)
```

## Re-process the two subsets

### CD3E high

I'll re-run the steps because of change in sample space; variable genes will be different (and are expected to be different)

```{r}
CD3E_high <- FindVariableFeatures(object = CD3E_high, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD3E_high) <-  VariableFeatures(CD3E_high) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(CD3E_high))]
VariableFeaturePlot(CD3E_high)
CD3E_high <- ScaleData(object = CD3E_high, features = VariableFeatures(object = CD3E_high), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD3E_high <- RunPCA(object = CD3E_high,
                features =  VariableFeatures(object = CD3E_high),
                dims = 1:30)
gc()
ElbowPlot(CD3E_high,ndims = 30)
CD3E_high <- RunHarmony(object = CD3E_high, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD3E_high <- RunUMAP(CD3E_high,dims = 1:20,reduction = "harmony")
CD3E_high <- RunTSNE(CD3E_high,dims = 1:20,reduction = "harmony")

CD3E_high <- FindNeighbors(CD3E_high, dims = 1:20,reduction = "harmony")
  
CD3E_high <- FindClusters(object = CD3E_high, resolution = 0.1)
gc()
```

```{r}
DimPlot(CD3E_high,reduction = "umap")
```
```{r}
DotPlot(CD3E_high,features = c("Cd3e","Cd4","Cd8a"))+RotatedAxis()
```

```{r}
save(CD3E_high,file = paste0(outdir,"CD3E_high.Robj"))
```

### CD3E low

```{r}
CD3E_low <- FindVariableFeatures(object = CD3E_low, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD3E_low) <-  VariableFeatures(CD3E_low) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(CD3E_low))]
VariableFeaturePlot(CD3E_low)
CD3E_low <- ScaleData(object = CD3E_low, features = VariableFeatures(object = CD3E_low), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD3E_low <- RunPCA(object = CD3E_low,
                features =  VariableFeatures(object = CD3E_low),
                dims = 1:30)
gc()
ElbowPlot(CD3E_low,ndims = 30)
CD3E_low <- RunHarmony(object = CD3E_low, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD3E_low <- RunUMAP(CD3E_low,dims = 1:20,reduction = "harmony")
CD3E_low <- RunTSNE(CD3E_low,dims = 1:20,reduction = "harmony")

CD3E_low <- FindNeighbors(CD3E_low, dims = 1:20,reduction = "harmony")
  
CD3E_low <- FindClusters(object = CD3E_low, resolution = 0.1)
gc()
```

```{r}
DimPlot(CD3E_low,reduction = "umap")

```


```{r}
DotPlot(CD3E_low,features = c("Cd3e","Cd4","Cd8a"))+RotatedAxis()
```

```{r}
save(CD3E_low,file = paste0(outdir,"CD3E_low.Robj"))
```

### Isolation and merging

```{r}
Cd3e_high_c2 <- subset(CD3E_high, idents = 2)
```

```{r}
table(Cd3e_high_c2$seurat_clusters)
```

```{r}
CD3E_high <- subset(CD3E_high, idents = c(0,1,3,4))
table(CD3E_high$seurat_clusters)
```

```{r}
Cd3e_low_high <- subset(CD3E_low,idents = c(2,4,5))
table(Cd3e_low_high$seurat_clusters)
```

```{r}
CD3E_low <- subset(CD3E_low,idents = c(0,1,3,6,7,8))
table(CD3E_low$seurat_clusters)
```

#### merge
```{r}
CD3E_high <- merge(CD3E_high,Cd3e_low_high)
CD3E_low <- merge(CD3E_low,Cd3e_high_c2)
```

## Process again

Note: previously merging will remove the previous processed layers. This does not seem to be the case for seurat 5

```{r}
CD3E_high <- FindVariableFeatures(object = CD3E_high, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD3E_high) <-  VariableFeatures(CD3E_high) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(CD3E_high))]
VariableFeaturePlot(CD3E_high)
CD3E_high <- ScaleData(object = CD3E_high, features = VariableFeatures(object = CD3E_high), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD3E_high <- RunPCA(object = CD3E_high,
                features =  VariableFeatures(object = CD3E_high),
                dims = 1:30)
gc()
ElbowPlot(CD3E_high,ndims = 30)
CD3E_high <- RunHarmony(object = CD3E_high, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD3E_high <- RunUMAP(CD3E_high,dims = 1:20,reduction = "harmony")
CD3E_high <- RunTSNE(CD3E_high,dims = 1:20,reduction = "harmony")

CD3E_high <- FindNeighbors(CD3E_high, dims = 1:20,reduction = "harmony")
  
CD3E_high <- FindClusters(object = CD3E_high, resolution = 0.1)
gc()
```

```{r}
DimPlot(CD3E_high,reduction = "umap")
DotPlot(CD3E_high,features = c("Cd3e","Cd4","Cd8a"))+RotatedAxis()
```
```{r}
save(CD3E_high,file = paste0(outdir,"CD3E_high.Robj"))
```

```{r}
CD3E_low <- FindVariableFeatures(object = CD3E_low, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD3E_low) <-  VariableFeatures(CD3E_low) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(CD3E_low))]
VariableFeaturePlot(CD3E_low)
CD3E_low <- ScaleData(object = CD3E_low, features = VariableFeatures(object = CD3E_low), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD3E_low <- RunPCA(object = CD3E_low,
                features =  VariableFeatures(object = CD3E_low),
                dims = 1:30)
gc()
ElbowPlot(CD3E_low,ndims = 30)
CD3E_low <- RunHarmony(object = CD3E_low, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD3E_low <- RunUMAP(CD3E_low,dims = 1:20,reduction = "harmony")
CD3E_low <- RunTSNE(CD3E_low,dims = 1:20,reduction = "harmony")

CD3E_low <- FindNeighbors(CD3E_low, dims = 1:20,reduction = "harmony")
  
CD3E_low <- FindClusters(object = CD3E_low, resolution = 0.1)
gc()
```


```{r}
DimPlot(CD3E_low,reduction = "umap")
DotPlot(CD3E_low,features = c("Cd3e","Cd4","Cd8a"))+RotatedAxis()
```

```{r}
save(CD3E_low,file = paste0(outdir,"CD3E_low.Robj"))
```


### merging and isolation again

```{r}
load(paste0(outdir,"CD3E_low.Robj"))
load(paste0(outdir,"CD3E_high.Robj"))
```

```{r}
DotPlot(CD3E_high,features = c("Cd3e","Cd4","Cd8a","Siglech","Ncr1","Adgre1","S100a8","Cd19"))+RotatedAxis()
DotPlot(CD3E_low,features = c("Cd3e","Cd4","Cd8a","Siglech","Ncr1","Adgre1","S100a8","Cd19"))+RotatedAxis()
```

```{r}
table(CD3E_high$seurat_clusters)
table(CD3E_low$seurat_clusters)
```



```{r}
FeaturePlot(CD3E_high,features = c("Cd3e","Adgre1"))
FeaturePlot(CD3E_high,features = c("Cd4"),label = T)
FeaturePlot(CD3E_high,features = c("Cd8a"),label = T)
FeaturePlot(CD3E_high,features = "nFeature_RNA",min.cutoff = 5000)
```

```{r}
FeaturePlot(CD3E_low,features = c("Cd3e","Adgre1"))
FeaturePlot(CD3E_low,features = "nFeature_RNA",min.cutoff = 5000)
```


```{r}
B_mixture <- subset(CD3E_high,idents = 7)
pDC <- subset(CD3E_high,idents = 5)
CD3E_high <- subset(CD3E_high, idents = seq(0,8,1)[!seq(0,8,1) %in% c(5,7)])
```

```{r}
NK <- subset(CD3E_low,idents = c(0,5))
CD3E_low <- subset(CD3E_low, idents = seq(0,6,1)[!seq(0,6,1) %in% c(0,5)])
```


Now, the previous CD3E_high is divided into three portions, and CD3E_low is divided to two portions. We want to look at each portion individually before merging them back.Instead of arbiturary doublet removal,removing transcriptionally vague clusters may be also a way.

We may want to see if we can further isolate mixtures.

```{r}
CD3E_high <- FindVariableFeatures(object = CD3E_high, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD3E_high) <-  VariableFeatures(CD3E_high) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(CD3E_high))]
VariableFeaturePlot(CD3E_high)
CD3E_high <- ScaleData(object = CD3E_high, features = VariableFeatures(object = CD3E_high), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD3E_high <- RunPCA(object = CD3E_high,
                features =  VariableFeatures(object = CD3E_high),
                dims = 1:30)
gc()
ElbowPlot(CD3E_high,ndims = 30)
CD3E_high <- RunHarmony(object = CD3E_high, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD3E_high <- RunUMAP(CD3E_high,dims = 1:20,reduction = "harmony")
CD3E_high <- RunTSNE(CD3E_high,dims = 1:20,reduction = "harmony")

CD3E_high <- FindNeighbors(CD3E_high, dims = 1:20,reduction = "harmony")
  
CD3E_high <- FindClusters(object = CD3E_high, resolution = 0.1)
gc()
```

```{r}
CD3E_low <- FindVariableFeatures(object = CD3E_low, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD3E_low) <-  VariableFeatures(CD3E_low) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(CD3E_low))]
VariableFeaturePlot(CD3E_low)
CD3E_low <- ScaleData(object = CD3E_low, features = VariableFeatures(object = CD3E_low), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD3E_low <- RunPCA(object = CD3E_low,
                features =  VariableFeatures(object = CD3E_low),
                dims = 1:30)
gc()
ElbowPlot(CD3E_low,ndims = 30)
CD3E_low <- RunHarmony(object = CD3E_low, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD3E_low <- RunUMAP(CD3E_low,dims = 1:20,reduction = "harmony")
CD3E_low <- RunTSNE(CD3E_low,dims = 1:20,reduction = "harmony")

CD3E_low <- FindNeighbors(CD3E_low, dims = 1:20,reduction = "harmony")
  
CD3E_low <- FindClusters(object = CD3E_low, resolution = 0.1)
gc()
```



```{r}
DotPlot(CD3E_high,features = c("Cd3e","Cd4","Cd8a","Siglech","Ncr1","Adgre1","S100a8","Cd19"))+RotatedAxis()
DotPlot(CD3E_low,features = c("Cd3e","Cd4","Cd8a","Siglech","Ncr1","Adgre1","S100a8","Cd19"))+RotatedAxis()
```

```{r}
CD3E_high_mixture <- subset(CD3E_high,idents = c(5,6))
NK_2 <- subset(CD3E_low,idents = c(3))
CD3E_high <- subset(CD3E_high, idents = seq(0,6,1)[!seq(0,6,1) %in% c(5,6)])
CD3E_low <- subset(CD3E_low, idents = seq(0,6,1)[!seq(0,6,1) %in% c(3)])
```

```{r}
B_mixture <- JoinLayers(B_mixture)
pDC <- JoinLayers(pDC)
CD3E_high_mixture <- JoinLayers(CD3E_high_mixture)
NK <- JoinLayers(NK)
NK_2 <- JoinLayers(NK_2)
```


Now merge the mixtures. Merge B_mixture, pDC and CD3E_high_mixture. Mixture two NK mixtures. 
```{r}

CD3E_high_mixture <- merge(CD3E_high_mixture , y = c(B_mixture,pDC))
NK_mixture <- merge(NK,NK_2)
```



