---
title: "0_prelimininary_processing"
output: html_document
date: "2024-08-06"
---

```{r}
library(Seurat)
library(tidyverse)
library(harmony)
library(DoubletFinder)
library(miQC)
library(SoupX)
library(data.table)
library(SeuratWrappers) #For miQC
```

T cell object generated was very preliminary and not the prerequisite object for the final T cell object.




# Single sample processing and merging without DoubletFinder

## Create Objects for each sample

```{r}
indir = "./alin_outs/" #Can be made into a command-line input
outdir = "./R_outs/"

```

```{r create object}
sample = "SR003099_NT-C"
NT_C <- CreateSeuratObject(Seurat::Read10X(paste(indir,"/", sample, "/outs/per_sample_outs/count/sample_filtered_feature_bc_matrix", sep = "")))
sample = "SR003099_NT-KO"
NT_KO <- CreateSeuratObject(Seurat::Read10X(paste(indir,"/", sample, "/outs/per_sample_outs/count/sample_filtered_feature_bc_matrix", sep = "")))
sample = "SR003099_T-C"
T_C <- CreateSeuratObject(Seurat::Read10X(paste(indir,"/", sample, "/outs/per_sample_outs/count/sample_filtered_feature_bc_matrix", sep = "")))
sample = "SR003099_T-KO"
T_KO <- CreateSeuratObject(Seurat::Read10X(paste(indir,"/", sample, "/outs/per_sample_outs/count/sample_filtered_feature_bc_matrix", sep = "")))
```

check initial cell number

```{r}
paste0("number of cells from NT-C is :", ncol(NT_C[["RNA"]]$counts))
paste0("number of genes from NT-C is :", nrow(NT_C[["RNA"]]$counts))
paste0("number of cells from NT-KO is :", ncol(NT_KO[["RNA"]]$counts))
paste0("number of genes from NT-KO is :", nrow(NT_KO[["RNA"]]$counts))
paste0("number of cells from T-C is :", ncol(T_C[["RNA"]]$counts))
paste0("number of genes from T-C is :", nrow(T_C[["RNA"]]$counts))
paste0("number of cells from T-KO is :", ncol(T_KO[["RNA"]]$counts))
paste0("number of genes from T-KO is :", nrow(T_KO[["RNA"]]$counts))
```

## Add additional metrics and perform FeatureScatter

```{r}
NT_C[["percent.mt"]] = PercentageFeatureSet(object = NT_C, pattern = "^MT-|^mt-")
NT_C$nFeature_RNA_log10 <- log10(NT_C$nFeature_RNA)
NT_C$nCount_RNA_log10 <- log10(NT_C$nCount_RNA)

NT_KO[["percent.mt"]] = PercentageFeatureSet(object = NT_KO, pattern = "^MT-|^mt-")
NT_KO$nFeature_RNA_log10 <- log10(NT_KO$nFeature_RNA)
NT_KO$nCount_RNA_log10 <- log10(NT_KO$nCount_RNA)

T_C[["percent.mt"]] = PercentageFeatureSet(object = T_C, pattern = "^MT-|^mt-")
T_C$nFeature_RNA_log10 <- log10(T_C$nFeature_RNA)
T_C$nCount_RNA_log10 <- log10(T_C$nCount_RNA)

T_KO[["percent.mt"]] = PercentageFeatureSet(object = T_KO, pattern = "^MT-|^mt-")
T_KO$nFeature_RNA_log10 <- log10(T_KO$nFeature_RNA)
T_KO$nCount_RNA_log10 <- log10(T_KO$nCount_RNA)
```


```{r}
FeatureScatter(NT_C, feature1 = "nFeature_RNA_log10", feature2 = "percent.mt")+
  ggtitle("NT-C")
FeatureScatter(NT_C, feature1 = "nFeature_RNA", feature2 = "percent.mt")+
  ggtitle("NT-C")
FeatureScatter(NT_KO, feature1 = "nFeature_RNA_log10", feature2 = "percent.mt")+
  ggtitle("NT-KO")
FeatureScatter(NT_KO, feature1 = "nFeature_RNA", feature2 = "percent.mt")+
  ggtitle("NT-KO")
FeatureScatter(T_C, feature1 = "nFeature_RNA_log10", feature2 = "percent.mt")+
  ggtitle("T-C")
FeatureScatter(T_C, feature1 = "nFeature_RNA", feature2 = "percent.mt")+
  ggtitle("T-C")
FeatureScatter(T_KO, feature1 = "nFeature_RNA_log10", feature2 = "percent.mt")+
  ggtitle("T-KO")
FeatureScatter(T_KO, feature1 = "nFeature_RNA", feature2 = "percent.mt")+
  ggtitle("T-KO")
```

## Run miQC for record keeping and checking threshold
Actual cutoff will be more conservative than miQC

### NT_C

```{r}
NT_C <- RunMiQC(NT_C, percent.mt = "percent.mt", nFeature_RNA = "nFeature_RNA", posterior.cutoff = 0.8,
               model.slot = "flexmix_model")
PlotMiQC(NT_C, color.by = "miQC.probability") + ggplot2::scale_color_gradient(low = "grey", high = "purple")+ggtitle("NT-C")
PlotMiQC(NT_C, color.by = "miQC.keep")+ggtitle("NT-C")
```
```{r}
table(NT_C$miQC.keep)
```


```{r}
Idents(NT_C) <- "NT-C"
VlnPlot(NT_C,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
```

```{r}
miQC_kept <- subset(NT_C, miQC.keep == "keep")
VlnPlot(miQC_kept,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
```

### NT_KO

```{r}
NT_KO <- RunMiQC(NT_KO, percent.mt = "percent.mt", nFeature_RNA = "nFeature_RNA", posterior.cutoff = 0.8,
               model.slot = "flexmix_model")
PlotMiQC(NT_KO, color.by = "miQC.probability") + ggplot2::scale_color_gradient(low = "grey", high = "purple")+ggtitle("NT-KO")
PlotMiQC(NT_KO, color.by = "miQC.keep")+ggtitle("NT-KO")
```
```{r}
table(NT_KO$miQC.keep)
```

```{r}
Idents(NT_KO) <- "NT-KO"
VlnPlot(NT_KO,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
```

```{r}
miQC_kept <- subset(NT_KO, miQC.keep == "keep")
Idents(miQC_kept) <- "NT-KO: miQC_kept"
VlnPlot(miQC_kept,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```

### T_C

```{r}
T_C <- RunMiQC(T_C, percent.mt = "percent.mt", nFeature_RNA = "nFeature_RNA", posterior.cutoff = 0.8,
               model.slot = "flexmix_model")
PlotMiQC(T_C, color.by = "miQC.probability") + ggplot2::scale_color_gradient(low = "grey", high = "purple")+ggtitle("T_C")
PlotMiQC(T_C, color.by = "miQC.keep")+ggtitle("T_C")
```

```{r}
table(T_C$miQC.keep)
```

```{r}
Idents(T_C) <- "T_C"
VlnPlot(T_C,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
```

```{r}
miQC_kept <- subset(T_C, miQC.keep == "keep")
Idents(miQC_kept) <- "T_C: miQC_kept"
VlnPlot(miQC_kept,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```

### T_KO

```{r}
T_KO <- RunMiQC(T_KO, percent.mt = "percent.mt", nFeature_RNA = "nFeature_RNA", posterior.cutoff = 0.8,
               model.slot = "flexmix_model")
PlotMiQC(T_KO, color.by = "miQC.probability") + ggplot2::scale_color_gradient(low = "grey", high = "purple")+ggtitle("T_KO")
PlotMiQC(T_KO, color.by = "miQC.keep")+ggtitle("T_KO")
```

```{r}
table(T_KO$miQC.keep)
```

```{r}
Idents(T_KO) <- "T_KO"
VlnPlot(T_KO,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
```

```{r}
miQC_kept <- subset(T_KO, miQC.keep == "keep")
Idents(miQC_kept) <- "T_KO: miQC_kept"
VlnPlot(miQC_kept,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```

## Trim cells by mt.percent first

### Try 5%

#### NT-C

```{r}
NT_C <- subset(NT_C, subset = percent.mt < 5)
paste0("The current number of cells from NT-C is :", ncol(NT_C[["RNA"]]$counts),". Was 2546")
VlnPlot(NT_C,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
```

```{r}
VlnPlot(NT_C,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```

#### NT-KO

```{r}
NT_KO <- subset(NT_KO, subset = percent.mt < 5)
paste0("The current number of cells from NT-KO is :", ncol(NT_KO[["RNA"]]$counts),". Was 3390")
VlnPlot(NT_KO,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
VlnPlot(NT_KO,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```
#### T-C


```{r}
T_C <- subset(T_C, subset = percent.mt < 5)
paste0("The current number of cells from T_C is :", ncol(T_C[["RNA"]]$counts),". Was 6919")
VlnPlot(T_C,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
VlnPlot(T_C,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```
#### T-KO


```{r}
T_KO <- subset(T_KO, subset = percent.mt < 5)
paste0("The current number of cells from T_C is :", ncol(T_KO[["RNA"]]$counts),". Was 8092")
VlnPlot(T_KO,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), split.by = "miQC.keep",split.plot = T, layer = "counts")
VlnPlot(T_KO,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"), layer = "counts")
```

## Processing single sample 

### NT-C

```{r}
NT_C <- NormalizeData(object = NT_C, normalization.method = "LogNormalize", scale.factor = 10000)
NT_C <- FindVariableFeatures(object = NT_C, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(NT_C) <-  VariableFeatures(NT_C)[!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(NT_C))]
VariableFeaturePlot(NT_C)
```

```{r}
NT_C <- ScaleData(object = NT_C, features = VariableFeatures(object = NT_C), vars.to.regress = c("nCount_RNA", "percent.mt"))
```

```{r}
NT_C <- RunPCA(NT_C,features =  VariableFeatures(object = NT_C))
NT_C <- RunUMAP(NT_C, dims = 1:10)
```

```{r}
FeaturePlot(NT_C,features = c("percent.mt","nFeature_RNA_log10"),reduction = "umap")
```


```{r}
FeaturePlot(NT_C,features = c("percent.mt","nCount_RNA_log10"),reduction = "umap")
```

```{r}
NT_C <- FindNeighbors(object = NT_C, dims = 1:10)              
NT_C <- FindClusters(object = NT_C, resolution = 0.1)
DimPlot(NT_C,reduction = "umap")
```
```{r eval=F}
save(NT_C, file = paste0(outdir,"NT_C.Robj"))
```

### NT-KO

```{r}
NT_KO <- NormalizeData(object = NT_KO, normalization.method = "LogNormalize", scale.factor = 10000)
NT_KO <- FindVariableFeatures(object = NT_KO, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(NT_KO) <-  VariableFeatures(NT_KO)[!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(NT_KO))]
VariableFeaturePlot(NT_KO)
```

```{r}
NT_KO <- ScaleData(object = NT_KO, features = VariableFeatures(object = NT_KO), vars.to.regress = c("nCount_RNA", "percent.mt"))
```

```{r}
NT_KO <- RunPCA(NT_KO,features =  VariableFeatures(object = NT_KO))
NT_KO <- RunUMAP(NT_KO, dims = 1:10)
```

```{r}
FeaturePlot(NT_KO,features = c("percent.mt","nFeature_RNA_log10"),reduction = "umap")
```

```{r}
FeaturePlot(NT_KO,features = c("percent.mt","nCount_RNA_log10"),reduction = "umap")
```

```{r}
NT_KO <- FindNeighbors(object = NT_KO, dims = 1:10)              
NT_KO <- FindClusters(object = NT_KO, resolution = 0.1)
DimPlot(NT_KO,reduction = "umap")
```

```{r eval=F}
save(NT_KO, file = paste0(outdir,"NT_KO.Robj"))
```

### T-C

```{r}
T_C <- NormalizeData(object = T_C, normalization.method = "LogNormalize", scale.factor = 10000)
T_C <- FindVariableFeatures(object = T_C, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(T_C) <-  VariableFeatures(T_C)[!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(T_C))]
VariableFeaturePlot(T_C)
```

```{r}
T_C <- ScaleData(object = T_C, features = VariableFeatures(object = T_C), vars.to.regress = c("nCount_RNA", "percent.mt"))
```

```{r}
T_C <- RunPCA(T_C,features =  VariableFeatures(object = T_C))
T_C <- RunUMAP(T_C, dims = 1:10)
```

```{r}
FeaturePlot(T_C,features = c("percent.mt","nFeature_RNA_log10"),reduction = "umap")
```

```{r}
FeaturePlot(T_C,features = c("percent.mt","nCount_RNA_log10"),reduction = "umap")
```

```{r}
T_C <- FindNeighbors(object = T_C, dims = 1:10)              
T_C <- FindClusters(object = T_C, resolution = 0.1)

```

```{r}
DimPlot(T_C,reduction = "umap")
```


```{r eval=F}
save(T_C, file = paste0(outdir,"T_C.Robj"))
```

### T-KO

```{r}
T_KO <- NormalizeData(object = T_KO, normalization.method = "LogNormalize", scale.factor = 10000)
T_KO <- FindVariableFeatures(object = T_KO, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(T_KO) <-  VariableFeatures(T_KO)[!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(T_KO))]
VariableFeaturePlot(T_KO)
```

```{r}
T_KO <- ScaleData(object = T_KO, features = VariableFeatures(object = T_KO), vars.to.regress = c("nCount_RNA", "percent.mt"))
```

```{r}
T_KO <- RunPCA(T_KO,features =  VariableFeatures(object = T_KO))
T_KO <- RunUMAP(T_KO, dims = 1:10)
```


```{r}
FeaturePlot(T_KO,features = c("percent.mt","nFeature_RNA_log10"),reduction = "umap")
```

```{r}
FeaturePlot(T_KO,features = c("percent.mt","nCount_RNA_log10"),reduction = "umap")
```

```{r}
T_KO <- FindNeighbors(object = T_KO, dims = 1:10)              
T_KO <- FindClusters(object = T_KO, resolution = 0.1)
DimPlot(T_KO,reduction = "umap")
```

```{r eval=F}
save(T_KO, file = paste0(outdir,"T_KO.Robj"))
```


## Merge into whole object

```{r}
whole <- merge(x=NT_C,y=c(NT_KO,T_C,T_KO),add.cell.ids=c("NT_C","NT_KO","T_C","T_KO"))
```


```{r }
NTC_range = range(grep("^NT_C",rownames(whole@meta.data)))

whole@meta.data$orig.ident[NTC_range[1]:NTC_range[2]] <- "NT_C"

NTKO_range = range(grep("^NT_KO",rownames(whole@meta.data)))

whole@meta.data$orig.ident[NTKO_range[1]:NTKO_range[2]] <- "NT_KO"

TC_range = range(grep("^T_C",rownames(whole@meta.data)))

whole@meta.data$orig.ident[TC_range[1]:TC_range[2]] <- "T_C"

TKO_range = range(grep("^T_KO",rownames(whole@meta.data)))

whole@meta.data$orig.ident[TKO_range[1]:TKO_range[2]] <- "T_KO"
```

```{r }
table(whole@meta.data$orig.ident)
```

```{r}
VlnPlot(whole,features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"),group.by = "orig.ident")
```


```{r eval=F}
save(whole,file = paste0(outdir,"whole.Robj"))
```


# Perform doublet finder and merge new version

---
title: "doublet removal"
output: html_document
---

```{r}
library(Seurat)
library(tidyverse)
library(harmony)
library(DoubletFinder)
library(miQC)
library(SoupX)
library(data.table)
library(SeuratWrappers) #For miQC
```

```{r}
indir = "./R_outs/"
outdir = "./R_outs/"
```

```{r load data}
load(paste0(outdir,"NT_C.Robj"))
load(paste0(outdir,"NT_KO.Robj"))
load(paste0(outdir,"T_C.Robj"))
load(paste0(outdir,"T_KO.Robj"))
```

```{r DoubletFinder functions}
paramSweep <- function(seu, PCs=1:10, sct = FALSE, num.cores=1) {
  require(Seurat); require(fields); require(parallel)
  ## Set pN-pK param sweep ranges
  pK <- c(0.0005, 0.001, 0.005, seq(0.01,0.3,by=0.01))
  pN <- seq(0.05,0.3,by=0.05)

  ## Remove pK values with too few cells
  min.cells <- round(nrow(seu@meta.data)/(1-0.05) - nrow(seu@meta.data))
  pK.test <- round(pK*min.cells)
  pK <- pK[which(pK.test >= 1)]

  ## Extract pre-processing parameters from original data analysis workflow
  orig.commands <- seu@commands

  ## Down-sample cells to 10000 (when applicable) for computational effiency
  if (nrow(seu@meta.data) > 10000) {
    real.cells <- rownames(seu@meta.data)[sample(1:nrow(seu@meta.data), 10000, replace=FALSE)]
    data <- seu[["RNA"]]$counts[ , real.cells]
    n.real.cells <- ncol(data)
  }

  if (nrow(seu@meta.data) <= 10000){
    real.cells <- rownames(seu@meta.data)
    data <- seu[["RNA"]]$counts
    n.real.cells <- ncol(data)
  }

  ## Iterate through pN, computing pANN vectors at varying pK
  #no_cores <- detectCores()-1
  if(num.cores>1){
    require(parallel)
    cl <- makeCluster(num.cores)
    output2 <- mclapply(as.list(1:length(pN)),
                        FUN = parallel_paramSweep,
                        n.real.cells,
                        real.cells,
                        pK,
                        pN,
                        data,
                        orig.commands,
                        PCs,
                        sct,mc.cores=num.cores)
    stopCluster(cl)
  }else{
    output2 <- lapply(as.list(1:length(pN)),
                      FUN = parallel_paramSweep,
                      n.real.cells,
                      real.cells,
                      pK,
                      pN,
                      data,
                      orig.commands,
                      PCs,
                      sct)
  }

  ## Write parallelized output into list
  sweep.res.list <- list()
  list.ind <- 0
  for(i in 1:length(output2)){
    for(j in 1:length(output2[[i]])){
      list.ind <- list.ind + 1
      sweep.res.list[[list.ind]] <- output2[[i]][[j]]
    }
  }

  ## Assign names to list of results
  name.vec <- NULL
  for (j in 1:length(pN)) {
    name.vec <- c(name.vec, paste("pN", pN[j], "pK", pK, sep = "_" ))
  }
  names(sweep.res.list) <- name.vec
  return(sweep.res.list)

}

doubletFinder <- function(seu, PCs, pN = 0.25, pK, nExp, reuse.pANN = FALSE, sct = FALSE, annotations = NULL) {
  require(Seurat); require(fields); require(KernSmooth)

  ## Generate new list of doublet classificatons from existing pANN vector to save time
  if (reuse.pANN != FALSE ) {
    pANN.old <- seu@meta.data[ , reuse.pANN]
    classifications <- rep("Singlet", length(pANN.old))
    classifications[order(pANN.old, decreasing=TRUE)[1:nExp]] <- "Doublet"
    seu@meta.data[, paste("DF.classifications",pN,pK,nExp,sep="_")] <- classifications
    return(seu)
  }

  if (reuse.pANN == FALSE) {
    ## Make merged real-artifical data
    real.cells <- rownames(seu@meta.data)
    data <- seu[["RNA"]]$counts[, real.cells]
    n_real.cells <- length(real.cells)
    n_doublets <- round(n_real.cells/(1 - pN) - n_real.cells)
    print(paste("Creating",n_doublets,"artificial doublets...",sep=" "))
    real.cells1 <- sample(real.cells, n_doublets, replace = TRUE)
    real.cells2 <- sample(real.cells, n_doublets, replace = TRUE)
    doublets <- (data[, real.cells1] + data[, real.cells2])/2
    colnames(doublets) <- paste("X", 1:n_doublets, sep = "")
    data_wdoublets <- cbind(data, doublets)
    # Keep track of the types of the simulated doublets
    if(!is.null(annotations)){
      stopifnot(typeof(annotations)=="character")
      stopifnot(length(annotations)==length(Cells(seu)))
      stopifnot(!any(is.na(annotations)))
      annotations <- factor(annotations)
      names(annotations) <- Cells(seu)
      doublet_types1 <- annotations[real.cells1]
      doublet_types2 <- annotations[real.cells2]
    }
    ## Store important pre-processing information
    orig.commands <- seu@commands

    ## Pre-process Seurat object
    if (sct == FALSE) {
      print("Creating Seurat object...")
      seu_wdoublets <- CreateSeuratObject(counts = data_wdoublets)

      print("Normalizing Seurat object...")
      seu_wdoublets <- NormalizeData(seu_wdoublets,
                                     normalization.method = orig.commands$NormalizeData.RNA@params$normalization.method,
                                     scale.factor = orig.commands$NormalizeData.RNA@params$scale.factor,
                                     margin = orig.commands$NormalizeData.RNA@params$margin)

      print("Finding variable genes...")
      seu_wdoublets <- FindVariableFeatures(seu_wdoublets,
                                            selection.method = orig.commands$FindVariableFeatures.RNA$selection.method,
                                            loess.span = orig.commands$FindVariableFeatures.RNA$loess.span,
                                            clip.max = orig.commands$FindVariableFeatures.RNA$clip.max,
                                            mean.function = orig.commands$FindVariableFeatures.RNA$mean.function,
                                            dispersion.function = orig.commands$FindVariableFeatures.RNA$dispersion.function,
                                            num.bin = orig.commands$FindVariableFeatures.RNA$num.bin,
                                            binning.method = orig.commands$FindVariableFeatures.RNA$binning.method,
                                            nfeatures = orig.commands$FindVariableFeatures.RNA$nfeatures,
                                            mean.cutoff = orig.commands$FindVariableFeatures.RNA$mean.cutoff,
                                            dispersion.cutoff = orig.commands$FindVariableFeatures.RNA$dispersion.cutoff)

      print("Scaling data...")
      seu_wdoublets <- ScaleData(seu_wdoublets,
                                 features = orig.commands$ScaleData.RNA$features,
                                 model.use = orig.commands$ScaleData.RNA$model.use,
                                 do.scale = orig.commands$ScaleData.RNA$do.scale,
                                 do.center = orig.commands$ScaleData.RNA$do.center,
                                 scale.max = orig.commands$ScaleData.RNA$scale.max,
                                 block.size = orig.commands$ScaleData.RNA$block.size,
                                 min.cells.to.block = orig.commands$ScaleData.RNA$min.cells.to.block)

      print("Running PCA...")
      seu_wdoublets <- RunPCA(seu_wdoublets,
                              features = orig.commands$ScaleData.RNA$features,
                              npcs = length(PCs),
                              rev.pca =  orig.commands$RunPCA.RNA$rev.pca,
                              weight.by.var = orig.commands$RunPCA.RNA$weight.by.var,
                              verbose=FALSE)
      pca.coord <- seu_wdoublets@reductions$pca@cell.embeddings[ , PCs]
      cell.names <- rownames(seu_wdoublets@meta.data)
      nCells <- length(cell.names)
      rm(seu_wdoublets); gc() # Free up memory
    }

    if (sct == TRUE) {
      require(sctransform)
      print("Creating Seurat object...")
      seu_wdoublets <- CreateSeuratObject(counts = data_wdoublets)

      print("Running SCTransform...")
      seu_wdoublets <- SCTransform(seu_wdoublets)

      print("Running PCA...")
      seu_wdoublets <- RunPCA(seu_wdoublets, npcs = length(PCs))
      pca.coord <- seu_wdoublets@reductions$pca@cell.embeddings[ , PCs]
      cell.names <- rownames(seu_wdoublets@meta.data)
      nCells <- length(cell.names)
      rm(seu_wdoublets); gc()
    }

    ## Compute PC distance matrix
    print("Calculating PC distance matrix...")
    dist.mat <- fields::rdist(pca.coord)

    ## Compute pANN
    print("Computing pANN...")
    pANN <- as.data.frame(matrix(0L, nrow = n_real.cells, ncol = 1))
    if(!is.null(annotations)){
      neighbor_types <- as.data.frame(matrix(0L, nrow = n_real.cells, ncol = length(levels(doublet_types1))))
    }
    rownames(pANN) <- real.cells
    colnames(pANN) <- "pANN"
    k <- round(nCells * pK)
    for (i in 1:n_real.cells) {
      neighbors <- order(dist.mat[, i])
      neighbors <- neighbors[2:(k + 1)]
      pANN$pANN[i] <- length(which(neighbors > n_real.cells))/k
      if(!is.null(annotations)){
        for(ct in unique(annotations)){
          neighbors_that_are_doublets = neighbors[neighbors>n_real.cells]
          if(length(neighbors_that_are_doublets) > 0){
            neighbor_types[i,] <-
              table( doublet_types1[neighbors_that_are_doublets - n_real.cells] ) +
              table( doublet_types2[neighbors_that_are_doublets - n_real.cells] )
            neighbor_types[i,] <- neighbor_types[i,] / sum( neighbor_types[i,] )
          } else {
            neighbor_types[i,] <- NA
          }
        }
      }
    }
    print("Classifying doublets..")
    classifications <- rep("Singlet",n_real.cells)
    classifications[order(pANN$pANN[1:n_real.cells], decreasing=TRUE)[1:nExp]] <- "Doublet"
    seu@meta.data[, paste("pANN",pN,pK,nExp,sep="_")] <- pANN[rownames(seu@meta.data), 1]
    seu@meta.data[, paste("DF.classifications",pN,pK,nExp,sep="_")] <- classifications
    if(!is.null(annotations)){
      colnames(neighbor_types) = levels(doublet_types1)
      for(ct in levels(doublet_types1)){
        seu@meta.data[, paste("DF.doublet.contributors",pN,pK,nExp,ct,sep="_")] <- neighbor_types[,ct]
      }
    }
    return(seu)
  }
}
parallel_paramSweep <- function(n, n.real.cells, real.cells, pK, pN, data, orig.commands, PCs, sct)  {

  sweep.res.list = list()
  list.ind = 0

  ## Make merged real-artifical data
  print(paste("Creating artificial doublets for pN = ", pN[n]*100,"%",sep=""))
  n_doublets <- round(n.real.cells/(1 - pN[n]) - n.real.cells)
  real.cells1 <- sample(real.cells, n_doublets, replace = TRUE)
  real.cells2 <- sample(real.cells, n_doublets, replace = TRUE)
  doublets <- (data[, real.cells1] + data[, real.cells2])/2
  colnames(doublets) <- paste("X", 1:n_doublets, sep = "")
  data_wdoublets <- cbind(data, doublets)

  ## Pre-process Seurat object
  if (sct == FALSE) {
    print("Creating Seurat object...")
    seu_wdoublets <- CreateSeuratObject(counts = data_wdoublets)

    print("Normalizing Seurat object...")
    seu_wdoublets <- NormalizeData(seu_wdoublets,
                                   normalization.method = orig.commands$NormalizeData.RNA@params$normalization.method,
                                   scale.factor = orig.commands$NormalizeData.RNA@params$scale.factor,
                                   margin = orig.commands$NormalizeData.RNA@params$margin)

    print("Finding variable genes...")
    seu_wdoublets <- FindVariableFeatures(seu_wdoublets,
                                          selection.method = orig.commands$FindVariableFeatures.RNA$selection.method,
                                          loess.span = orig.commands$FindVariableFeatures.RNA$loess.span,
                                          clip.max = orig.commands$FindVariableFeatures.RNA$clip.max,
                                          mean.function = orig.commands$FindVariableFeatures.RNA$mean.function,
                                          dispersion.function = orig.commands$FindVariableFeatures.RNA$dispersion.function,
                                          num.bin = orig.commands$FindVariableFeatures.RNA$num.bin,
                                          binning.method = orig.commands$FindVariableFeatures.RNA$binning.method,
                                          nfeatures = orig.commands$FindVariableFeatures.RNA$nfeatures,
                                          mean.cutoff = orig.commands$FindVariableFeatures.RNA$mean.cutoff,
                                          dispersion.cutoff = orig.commands$FindVariableFeatures.RNA$dispersion.cutoff)

    print("Scaling data...")
    seu_wdoublets <- ScaleData(seu_wdoublets,
                               features = orig.commands$ScaleData.RNA$features,
                               model.use = orig.commands$ScaleData.RNA$model.use,
                               do.scale = orig.commands$ScaleData.RNA$do.scale,
                               do.center = orig.commands$ScaleData.RNA$do.center,
                               scale.max = orig.commands$ScaleData.RNA$scale.max,
                               block.size = orig.commands$ScaleData.RNA$block.size,
                               min.cells.to.block = orig.commands$ScaleData.RNA$min.cells.to.block)

    print("Running PCA...")
    seu_wdoublets <- RunPCA(seu_wdoublets,
                            features = orig.commands$ScaleData.RNA$features,
                            npcs = length(PCs),
                            rev.pca =  orig.commands$RunPCA.RNA$rev.pca,
                            weight.by.var = orig.commands$RunPCA.RNA$weight.by.var,
                            verbose=FALSE)
  }

  if (sct == TRUE) {
    require(sctransform)
    print("Creating Seurat object...")
    seu_wdoublets <- CreateSeuratObject(counts = data_wdoublets)

    print("Running SCTransform...")
    seu_wdoublets <- SCTransform(seu_wdoublets)

    print("Running PCA...")
    seu_wdoublets <- RunPCA(seu_wdoublets, npcs = length(PCs))
  }

  ## Compute PC distance matrix
  print("Calculating PC distance matrix...")
  nCells <- nrow(seu_wdoublets@meta.data)
  pca.coord <- seu_wdoublets@reductions$pca@cell.embeddings[ , PCs]
  rm(seu_wdoublets)
  gc()
  dist.mat <- fields::rdist(pca.coord)[,1:n.real.cells]

  ## Pre-order PC distance matrix prior to iterating across pK for pANN computations
  print("Defining neighborhoods...")
  for (i in 1:n.real.cells) {
    dist.mat[,i] <- order(dist.mat[,i])
  }

  ## Trim PC distance matrix for faster manipulations
  ind <- round(nCells * max(pK))+5
  dist.mat <- dist.mat[1:ind, ]

  ## Compute pANN across pK sweep
  print("Computing pANN across all pK...")
  for (k in 1:length(pK)) {
    print(paste("pK = ", pK[k], "...", sep = ""))
    pk.temp <- round(nCells * pK[k])
    pANN <- as.data.frame(matrix(0L, nrow = n.real.cells, ncol = 1))
    colnames(pANN) <- "pANN"
    rownames(pANN) <- real.cells
    list.ind <- list.ind + 1

    for (i in 1:n.real.cells) {
      neighbors <- dist.mat[2:(pk.temp + 1),i]
      pANN$pANN[i] <- length(which(neighbors > n.real.cells))/pk.temp
    }

    sweep.res.list[[list.ind]] <- pANN

  }

  return(sweep.res.list)
}
```

## NT_C

```{r NT_C sweep}
sweep.res.list <- paramSweep(NT_C, PCs = 1:30, sct = FALSE)


sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)




bcmvn <- find.pK(sweep.stats)


homotypic.prop <- modelHomotypic(NT_C@meta.data$seurat_clusters)          
nExp_poi <- round(0.05*nrow( NT_C@meta.data))  ## Assuming 5% doublet formation rate 
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))

```

```{r}
 NT_C <- doubletFinder( NT_C, PCs = 1:30, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)

head(NT_C@meta.data) # This is too much for whole


```

```{r}
FeaturePlot( NT_C,features = "pANN_0.25_0.09_122",reduction = "umap")
DimPlot(NT_C,reduction = "umap",group.by = "DF.classifications_0.25_0.09_122")
FeaturePlot( NT_C,features = "Cd4",reduction = "umap")
FeaturePlot( NT_C,features = "Cd8a",reduction = "umap")
```
```{r}
NT_C <- subset(NT_C,subset = DF.classifications_0.25_0.09_122 == "Singlet")
```

```{r}
save(NT_C,file = paste0(outdir,"NT_C_doubRemoved.Robj"))
```

## NT_KO

```{r NT_KO sweep}
sweep.res.list <- paramSweep(NT_KO, PCs = 1:30, sct = FALSE)


sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)




bcmvn <- find.pK(sweep.stats)


homotypic.prop <- modelHomotypic(NT_KO@meta.data$seurat_clusters)          
nExp_poi <- round(0.05*nrow( NT_KO@meta.data))  ## Assuming 5% doublet formation rate 
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))

NT_KO <- doubletFinder( NT_KO, PCs = 1:30, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)

head(NT_KO@meta.data) # This is too much for whole


```


```{r}
FeaturePlot( NT_KO,features = "pANN_0.25_0.09_163",reduction = "umap")
DimPlot(NT_KO,reduction = "umap",group.by = "DF.classifications_0.25_0.09_163")
FeaturePlot( NT_KO,features = "Cd4",reduction = "umap")
FeaturePlot( NT_KO,features = "Cd8a",reduction = "umap")
```

```{r}
NT_KO <- subset(NT_KO,subset = DF.classifications_0.25_0.09_163 == "Singlet")
```

```{r}
save(NT_KO,file = paste0(outdir,"NT_KO_doubRemoved.Robj"))
```


## T_C

```{r}
sweep.res.list <- paramSweep(T_C, PCs = 1:30, sct = FALSE)


sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)




bcmvn <- find.pK(sweep.stats)


homotypic.prop <- modelHomotypic(T_C@meta.data$seurat_clusters)          
nExp_poi <- round(0.05*nrow(T_C@meta.data))  ## Assuming 5% doublet formation rate 
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))

T_C <- doubletFinder(T_C, PCs = 1:30, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)

head(T_C@meta.data) # This is too much for whole

```

```{r}
FeaturePlot(T_C,features = "pANN_0.25_0.09_328",reduction = "umap")
DimPlot(T_C,reduction = "umap",group.by = "DF.classifications_0.25_0.09_328")
FeaturePlot(T_C,features = "Cd4",reduction = "umap")
FeaturePlot(T_C,features = "Cd8a",reduction = "umap")
```

```{r}
T_C <- subset(T_C,subset = DF.classifications_0.25_0.09_328 == "Singlet")
```

```{r}
save(T_C,file = paste0(outdir,"T_C_doubRemoved.Robj"))
```




## T_KO

```{r}
sweep.res.list <- paramSweep(T_KO, PCs = 1:30, sct = FALSE)


sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)




bcmvn <- find.pK(sweep.stats)


homotypic.prop <- modelHomotypic(T_KO@meta.data$seurat_clusters)          
nExp_poi <- round(0.05*nrow( T_KO@meta.data))  ## Assuming 5% doublet formation rate 
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))

T_KO <- doubletFinder( T_KO, PCs = 1:30, pN = 0.25, pK = 0.09, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)

head(T_KO@meta.data) # This is too much for whole

```

```{r}
FeaturePlot( T_KO,features = "pANN_0.25_0.09_384",reduction = "umap")
DimPlot(T_KO,reduction = "umap",group.by = "DF.classifications_0.25_0.09_384")
FeaturePlot( T_KO,features = "Cd4",reduction = "umap")
FeaturePlot( T_KO,features = "Cd8a",reduction = "umap")
```

```{r}
T_KO <- subset(T_KO,subset = DF.classifications_0.25_0.09_384 == "Singlet")
```

```{r}
save(T_KO,file = paste0(outdir,"T_KO_doubRemoved.Robj"))
```



## merge and analyze

```{r}
whole <- merge(x=NT_C,y=c(NT_KO,T_C,T_KO),add.cell.ids=c("NT_C","NT_KO","T_C","T_KO"))
```

```{r }
NTC_range = range(grep("^NT_C",rownames(whole@meta.data)))

whole@meta.data$orig.ident[NTC_range[1]:NTC_range[2]] <- "NT_C"

NTKO_range = range(grep("^NT_KO",rownames(whole@meta.data)))

whole@meta.data$orig.ident[NTKO_range[1]:NTKO_range[2]] <- "NT_KO"

TC_range = range(grep("^T_C",rownames(whole@meta.data)))

whole@meta.data$orig.ident[TC_range[1]:TC_range[2]] <- "T_C"

TKO_range = range(grep("^T_KO",rownames(whole@meta.data)))

whole@meta.data$orig.ident[TKO_range[1]:TKO_range[2]] <- "T_KO"
```


```{r}
whole$DF.classifications_0.25_0.09_122 <- NULL
whole$DF.classifications_0.25_0.09_163 <- NULL
whole$DF.classifications_0.25_0.09_328 <- NULL
whole$DF.classifications_0.25_0.09_384 <- NULL
```

```{r }
table(whole@meta.data$orig.ident)
```

```{r eval=F}
save(whole,file = paste0(outdir,"whole_DoubletFinder.Robj"))
```

```{r}
whole <- NormalizeData(object = whole, normalization.method = "LogNormalize", scale.factor = 10000)
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-A", VariableFeatures(whole))]
VariableFeaturePlot(whole)
whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA", "percent.mt"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```

```{r}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
whole <- RunUMAP(whole,dims = 1:40,reduction = "harmony")
whole <- RunTSNE(whole,dims = 1:40,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:40,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r eval=F}
save(whole,file = paste0(outdir,"whole_DoubletFinder.Robj"))
```

```{r}
load(paste0(outdir,"whole_DoubletFinder.Robj"))
```

```{r}
DimPlot(whole, reduction = "umap",label = T)
```

```{r}
DimPlot(whole, reduction = "umap",label = T,split.by = "orig.ident",ncol = 2)

```

```{r}
DimPlot(whole, reduction = "umap",group.by = "RNA_snn_res.0.1",label = T)
```


```{r}
DimPlot(whole, reduction = "umap",group.by = "orig.ident",shuffle = T)
DimPlot(whole, reduction = "umap",group.by = "RNA_snn_res.0.6",label = T)
DimPlot(whole, reduction = "umap")
DimPlot(whole, reduction = "tsne")
FeaturePlot(whole,features = c("Cd8a"))
FeaturePlot(whole,features = c("Cd19"))
DimPlot(whole, reduction = "umap",group.by = "miQC.keep",shuffle = T)
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"))+RotatedAxis()
```

```{r}
VlnPlot(whole, features = c("percent.mt","nFeature_RNA_log10","nCount_RNA_log10"),group.by = "orig.ident")
```

```{r}
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Lyz2","Fn1","Xcr1","P2ry12","Ncr1","Siglech","S100a8","Mpo","Fcer2a","Flt3","Cd200r3","Gata3","Il7r","Mki67"),group.by = "RNA_snn_res.0.6")+RotatedAxis()
```

```{r}
VlnPlot(whole, features = c("Cd3e"),group.by = "RNA_snn_res.0.6")
Idents(whole) <- whole$RNA_snn_res.0.6
FeaturePlot(whole,features = c("Cd3e"),label = T)
```

```{r}
VlnPlot(whole, features = c("Cd19"),group.by = "RNA_snn_res.0.6")
Idents(whole) <- whole$RNA_snn_res.0.6
FeaturePlot(whole,features = c("Cd19"),label = T)
```
```{r}
VlnPlot(whole, features = c("Cd8a"),group.by = "RNA_snn_res.0.6")
Idents(whole) <- whole$RNA_snn_res.0.6
FeaturePlot(whole,features = c("Cd8a"),label = T)
```

```{r}
VlnPlot(whole, features = c("Ncr1"),group.by = "RNA_snn_res.0.1")
Idents(whole) <- whole$RNA_snn_res.0.1
FeaturePlot(whole,features = c("Ncr1"),label = T)
```

```{r}
cluster_prop <- whole@meta.data %>% group_by(orig.ident,RNA_snn_res.0.6) %>%
  summarise(n = n()) %>% mutate(proportion = 100*n/sum(n))

p1 <- ggplot(cluster_prop,aes(x = orig.ident, y = proportion,fill = orig.ident))+
  geom_bar(stat = "identity", position = 'dodge')+
  facet_wrap(~RNA_snn_res.0.6,scales = "free")+guides(fill=guide_legend(title="Condition"))+ ylab("Percentage of all cells")+
  theme(
  plot.title = element_blank(),
  axis.title.x = element_blank())+RotatedAxis()
p1
g <- gridExtra::grid.arrange(grobs = list(p1),nrow = 2)
ggsave(filename = paste0(outdir,"Proportion_barplots_res06_whole_DF_V2.pdf"), plot = g, units = "cm",height = 30,width = 20)
```


```{r}
T_cells <- subset(whole, idents = c(0,1,2,4,5,7,8,12,13,17))
save(T_cells,file = paste0(outdir,"T_cells_DF_res06.Robj"))
```


# **Archived** version of second round QC:


```{r}
library(Seurat)
library(tidyverse)
library(harmony)
```

```{r}
indir = "./R_outs/" #Can be made into a command-line input
outdir = "./R_outs/"
```

```{r}
load(paste0(outdir,"whole.Robj"))
```

## run the usual pipeline

```{r}
table(whole@meta.data$orig.ident)
```

```{r}
ggplot(whole@meta.data)+
  geom_histogram(aes(x=percent.mt,y = after_stat(width*density*100)),bins = 100)+
  facet_wrap(vars(orig.ident),ncol=1)+
  geom_vline(xintercept = 5,col="red") +
  ylab("proportion")+
  xlab("percentage of mitochondrial RNA")
```

```{r}
ggplot(whole@meta.data, aes(x = log10(nFeature_RNA))) +
  geom_histogram(bins = 100, aes(y = after_stat(width*density*100))) +
  theme_classic() +
  facet_wrap(~orig.ident, ncol = 1) +
  geom_vline(xintercept = 2.5,col="red") + 
  geom_vline(xintercept = 4,col="red") + 
  ggtitle("nGene") +
  ylab("")
```

```{r}
ggplot(whole@meta.data, aes(x = log10(nCount_RNA))) +
  geom_histogram(bins = 100, aes(y = after_stat(width*density*100))) +
  theme_classic() +
  facet_wrap(~orig.ident, ncol = 1) +
  geom_vline(xintercept = 2.5,col="red") + 
  geom_vline(xintercept = 4.5,col="red") + 
  ggtitle("nCount") +
  ylab("")
```


```{r}
#whole <- NormalizeData(object = whole, normalization.method = "LogNormalize", scale.factor = 10000)
#there is no need to normalize again; normalization is cell specific
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^TRA|^TRB|^IGH|^IGK|MAMU-A", VariableFeatures(whole))]
VariableFeaturePlot(whole)
```


```{r}
whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA", "percent.mt"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```

```{r}
DimPlot(whole, reduction = "pca",group.by = "orig.ident",dims = c(3,4),shuffle = T)
```


```{r}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
whole <- RunUMAP(whole,dims = 1:20,reduction = "harmony")
whole <- RunTSNE(whole,dims = 1:20,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```


```{r}
DimPlot(whole, reduction = "umap",group.by = "orig.ident",shuffle = T)
```

```{r}
DimPlot(whole, reduction = "umap",group.by = "RNA_snn_res.0.1")
```

```{r}
FeaturePlot(whole,features = c("Cd8a"))
```

```{r}
DimPlot(whole, reduction = "umap")
```


```{r}
FeaturePlot(whole,features = c("Cd8a"))
```

```{r}
DimPlot(whole, reduction = "umap",group.by = "miQC.keep",shuffle = T)
```


```{r}
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"),group.by = "RNA_snn_res.0.1")+RotatedAxis()
```
```{r}
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"),group.by = "orig.ident")+RotatedAxis()
```

```{r}
bulk <- AggregateExpression(whole,group.by = "orig.ident", return.seurat = TRUE)
Cells(bulk)
```

```{r}
DotPlot(bulk,features = c("Cd3e","Cd8a","Cd4","Cd19","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"))+RotatedAxis()
```
```{r}
whole <- JoinLayers(whole) #This is seurat 5 specific
```


```{r eval=F}
save(whole,file = paste0(outdir,"whole_processed_wHarmony.Robj"))
```

## Jan 5 processing by further trimming down to miQC.keep


```{r}
whole <- subset(whole, subset = miQC.keep=="keep")
```


```{r}
table(whole$miQC.keep)
```


```{r}
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|MAMU-A", VariableFeatures(whole))]
VariableFeaturePlot(whole)

whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA", "percent.mt"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```

```{r}
DimPlot(whole, reduction = "pca",group.by = "orig.ident",dims = c(3,4),shuffle = T)
```


```{r}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
whole <- RunUMAP(whole,dims = 1:20,reduction = "harmony")
whole <- RunTSNE(whole,dims = 1:20,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r}
DimPlot(whole, reduction = "umap",group.by = "orig.ident",shuffle = T)
DimPlot(whole, reduction = "umap",group.by = "RNA_snn_res.0.1")
DimPlot(whole, reduction = "umap")
```

```{r}
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"),group.by = "RNA_snn_res.0.1")+RotatedAxis()
DotPlot(whole,features = c("Cd3e","Cd8a","Cd4","Cd19","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Ccr2","Fcer1a","Fcer2a","Flt3"),group.by = "orig.ident")+RotatedAxis()
```

```{r}
save(whole, file = paste0(outdir,"whole_processed_wHarmony_miQC.Robj"))
```

