---
title: "Pub6_re-do_TCR_analysis_and_plots"
output: html_document
date: "2024-09-23"
---

```{r}
library(Seurat)
library(tidyverse)
library(scater)
library(data.table)
library(RColorBrewer)
library(gridExtra)
library(ggrepel)
library(kableExtra)
library(scRepertoire)
library(ggraph)
library(circlize)
library(scales)
```


```{r}
pal <- c(brewer.pal(n = 12,name = "Set3"),brewer.pal(n = 12,name = "Paired"),brewer.pal(n = 8,name = "Accent"))
pal[2] <- "darkred"
pal[9] <- "grey25"
pal[23] <- "turquoise4"
pal[19] <- colors()[60]
pal[13] <- "darkblue"
pal[28] <- colors()[16]
pal[31] <- colors()[54] 
pal <- c(pal,colors()[c(85,84,83,89,45,31,41,42,26,29,10,139,107,108,120,109,119,121,143)])
```


```{r}
indir <- "../V2_TCR_analysis_and_V3_Update/VDJ/"
RDSdir <- "../V2_processing_analysis/R_outs/"
outdir <- "TCR_outs_Sep23/"
if (!dir.exists(outdir)) dir.create(outdir)
```

```{r}
load(paste0(RDSdir,"lymphoid_convT_V3p3_merged_scRep_aa.Robj"))
```

## prepare TCR-seq analysis object

```{r}
samples <- list.files(path = indir)
```

```{r}
contig_files <- paste(indir,samples,"filtered_contig_annotations.csv",sep = "/")
```

```{r}
contig_list <- lapply(as.list(contig_files), read.csv)
names(contig_list) <- samples
```

```{r}
combined <- combineTCR(contig_list,
                       removeNA = T,
                       removeMulti = T,
                        samples = samples)
```

```{r}
relative_abundance <- as.data.frame(clonalHomeostasis(combined, 
                  cloneCall = "aa",exportTable = T))
relative_abundance = relative_abundance[,colSums(relative_abundance) > 0] 
relative_abundance$condition = rownames(relative_abundance)

relative_abundance = reshape2::melt(relative_abundance, id = "condition",variable.name = "ClonalType", value.name = "Proportion")
g <- ggplot(relative_abundance)+
  geom_bar(aes(x = condition, y = Proportion, fill = ClonalType ),stat = "identity",position = "stack")+scale_fill_manual(values = c("Small (1e-04 < X <= 0.001)" = "#56106EFF", "Medium (0.001 < X <= 0.01)" = "#BB3754FF", "Large (0.01 < X <= 0.1)" = "#F98C0AFF"))+theme_classic()
g
ggsave(filename = paste0(outdir,"Fig4F_relative_clone_abundance_aa.pdf"),plot = g, width = 10,height = 8)
```

```{r}
write.table(relative_abundance,file = paste0(outdir,"/scRep_relative_abundance_aa_table.tsv"),sep = "\t",col.names = T,row.names = F)
```


## prepare merging into seurat object

```{r}
colnames(screp.aa@meta.data)
```
clean up previous cloumns

```{r}
screp.aa$CTaa <- NULL
screp.aa$CTgene <- NULL
screp.aa$CTnt <- NULL
screp.aa$CTstrict <- NULL
screp.aa$Frequency <- NULL
screp.aa$cloneType <- NULL
screp.aa$clonalProportion <- NULL
screp.aa$clonalFrequency <- NULL
screp.aa$cloneSize <- NULL
```

```{r}
colnames(screp.aa@meta.data)
```
```{r}
screp.aa <- combineExpression(combined, 
                         screp.aa, 
                         cloneCall="aa", 
                         group.by = "sample", 
                         proportion = TRUE)
```

```{r}
colnames(screp.aa@meta.data)
```

```{r}
table(screp.aa$cloneSize)
```

## multimodal analysis

### round1 top clone sanity check

```{r}
top_clone_rank <- screp.aa@meta.data %>% filter((!is.na(CTaa)) & (cloneSize == "Large (0.01 < X <= 0.1)" | cloneSize == "Medium (0.001 < X <= 0.01)")) %>% group_by(orig.ident,celltype,CTaa,clonalProportion,cloneSize) %>% summarise() %>% ungroup() %>% slice_max(order_by = clonalProportion, n = 30) %>% ungroup()
```

```{r}
write.table(top_clone_rank, file = paste0(outdir,"/Top_clone_aa_without_grouping.tsv"),sep = "\t",col.names = T,row.names = F)
```

```{r}
g <- ggplot(top_clone_rank) + 
  geom_bar(aes(x = reorder(CTaa, -clonalProportion), y = clonalProportion, fill = orig.ident),stat = "identity",position = "dodge")+
  facet_wrap(~orig.ident, scale = "free")+theme_classic()+
  xlab("Clonotype by amino acid")+
  theme(axis.text.x = element_text(angle = 90,size = 5))
g
```


```{r}
top_clone_rank <- screp.aa@meta.data %>% filter((!is.na(CTaa)) & (cloneSize == "Large (0.01 < X <= 0.1)" | cloneSize == "Medium (0.001 < X <= 0.01)")) %>% group_by(orig.ident,celltype,CTaa,clonalProportion,cloneSize) %>% summarise() %>% ungroup() %>% filter(celltype != "Proliferating T") %>% group_by(orig.ident) %>% slice_max(order_by = clonalProportion, n = 10) %>% ungroup()
```

```{r}
g <- ggplot(top_clone_rank) + 
  geom_bar(aes(x = reorder(CTaa, -clonalProportion), y = clonalProportion, fill = orig.ident),stat = "identity",position = "dodge")+
  facet_grid(~celltype, scale = "free")+theme_classic()+
  xlab("Clonotype by amino acid")+
  theme(axis.text.x = element_text(angle = 90,size = 8))
g
ggsave(paste0(outdir,"FigS10C_CD3r_chain_aa_top10_from_each_sample_by_prop.pdf"),plot = g, width = 20, height = 15)
```

```{r}
write.table(top_clone_rank, file = paste0(outdir,"/Top_10_clone_aa_by_sample_wo_prolif.tsv"),sep = "\t",col.names = T,row.names = F)
```

The cell type information has been already updated (for one NKT-like clone and one CD4-like clone), therefore I'll perform some additional sanity check.

NKT clone: "CAVMSNYNVLYF_CALGNTEVFF"
top 1 CD4 clone: "CAMREGSYTNKVVF_CASSPGQGYERLFF"

Previous changes:

```
screp.aa@meta.data <- screp.aa@meta.data %>% mutate(
  celltype = case_when(
    highlight == "CAVMSNYNVLYF_CALGNTEVFF" & (celltype == "CD8T" | celltype == "CD4T") ~ "NKT",
    .default = celltype
  ),
  CD4_top1_troubleshoot = case_match(
    CTaa,
    "CAMREGSYTNKVVF_CASSPGQGYERLFF" ~ "CAMREGSYTNKVVF_CASSPGQGYERLFF",
    .default = "Other"
  )
)
```

cells with CAVMSNYNVLYF_CALGNTEVFF that were labeled as CD8T or CD4T were corrected to NKT


```
screp.aa@meta.data <- screp.aa@meta.data %>% mutate(
  celltype = case_when(
    highlight == "CAVMSNYNVLYF_CALGNTEVFF" & (celltype == "CD8T" | celltype == "CD4T") ~ "NKT",
    CTaa == "CAMREGSYTNKVVF_CASSPGQGYERLFF" & celltype != "Proliferating T" ~ "CD4T",
    .default = celltype
  )
)
```

```{r}
VlnPlot(screp.aa,group.by = "CD4_top1_troubleshoot",features = "Cd4")
VlnPlot(screp.aa,group.by = "CD4_top1_troubleshoot",features = "Cd8a")
VlnPlot(screp.aa,group.by = "CD4_top1_troubleshoot",features = "Klrb1c")
```

### check top clone location

```{r}
meta <- cbind(screp.aa@meta.data,screp.aa@reductions$umap@cell.embeddings)
```


```{r}
meta <- meta %>% 
  mutate(clone_type = case_when(
    CTaa %in% top_clone_rank$CTaa ~ "Top",
    .default = "non-Top"
  ))
```


```{r}
g <- ggplot(meta)+
  geom_point(aes(x= umap_1, y = umap_2, color = clone_type,size = clonalProportion))+
  facet_wrap(~orig.ident)+scale_color_manual(values = c("grey50","red2"))+scale_radius(range = c(0,1.5))+theme_classic()
g
# ggsave(paste0(outdir,"Fig4_top_clones_on_umap_by_geno.pdf"),plot = g, width = 10, height = 8)
```

### Gini Index


```{r}
library(DescTools)
```


```{r}
Gini_df  <- screp.aa@meta.data %>% filter((!is.na(cloneSize))) %>% group_by(orig.ident,celltype,CTaa,clonalFrequency) %>% summarise(cloneFreq = n()) %>% ungroup() %>% group_by(orig.ident,celltype) %>% mutate(Gini_index = Gini(clonalFrequency)) %>% ungroup()
```

```{r}
g <- ggplot(Gini_df) +
  geom_bar(aes(x = orig.ident, y = Gini_index, fill = orig.ident),stat = "identity",position = "dodge")+
  facet_wrap(~celltype,scales = "free")+xlab("Condition")+theme_classic()+ylab("Gini Index")
g
ggsave(filename = paste0(outdir,"/Fig4_Gini_index_plot.pdf"),width = 10,height = 6)
g <- ggplot(Gini_df) +
  geom_bar(aes(x = orig.ident, y = Gini_index, fill = orig.ident),stat = "identity",position = "dodge")+xlab("Condition")+theme_classic()+ylab("Gini Index")
g
ggsave(filename = paste0(outdir,"/Fig4_Gini_index_plot_bulk.pdf"),width = 10,height = 6)
```

### sanity check clones

```{r}
NKT_clones <- meta %>% filter(celltype == "NKT") %>% group_by(CTaa, clonalProportion,clonalFrequency,orig.ident) %>% summarise()
```


```{r}
meta <- meta %>% 
  mutate(clone_type = case_when(
    CTaa %in% NKT_clones$CTaa ~ "NKT_clone",
    .default = "Other"
  ))
```


```{r}
g <- ggplot(meta)+
  geom_point(aes(x= umap_1, y = umap_2, color = clone_type,size = clonalProportion))+
  facet_wrap(~orig.ident)+scale_color_manual(values = c("red2","grey50"))+scale_radius(range = c(0,1.5))+theme_classic()
g
# ggsave(paste0(outdir,"Fig4_top_clones_on_umap_by_geno.pdf"),plot = g, width = 10, height = 8)
```

```{r}
colnames(meta)
```


```{r}
g <- ggplot(meta)+
  geom_point(aes(x= umap_1, y = umap_2, color = highlight,size = clonalProportion))+
  facet_wrap(~orig.ident)+scale_color_manual(values = c("red2","blue2","grey50"))+scale_radius(range = c(0,1.5))+theme_classic()
g
```
```{r}
NKT_clones %>% arrange(-clonalProportion)
```


```{r}
screp.aa@meta.data <- screp.aa@meta.data %>% mutate(
  Clone_troubleshoot = case_match(
    CTaa,
    "CAVMSNYNVLYF_CALGNTEVFF" ~ "NKT1: CAVMSNYNVLYF_CALGNTEVFF",
    "CALSTPYSNNRLTL_CASSLGTGEVFF" ~ "NKT2: CALSTPYSNNRLTL_CASSLGTGEVFF",
    "CAVTNSNNRIFF_CAWSLSWGGSAETLYF" ~ "NKT3: CAVTNSNNRIFF_CAWSLSWGGSAETLYF",
    "CAVNNYAQGLTF_CASSPPTGEDTQYF" ~ "NKT4: CAVNNYAQGLTF_CASSPPTGEDTQYF",
    "CAASERADYANKMIF_CASSRPGQGAQAPLF" ~ "NKT5: CAASERADYANKMIF_CASSRPGQGAQAPLF",
    "CAVRPSNTNKVVF_CASSPNTEVFF" ~ "Mix: CAVRPSNTNKVVF_CASSPNTEVFF",
    .default = "Other"
  )
)
```


```{r}
VlnPlot(screp.aa,group.by = "Clone_troubleshoot",features = "Cd4")
VlnPlot(screp.aa,group.by = "Clone_troubleshoot",features = "Cd8a")
VlnPlot(screp.aa,group.by = "Clone_troubleshoot",features = "Klrb1c")
VlnPlot(screp.aa,group.by = "Clone_troubleshoot",features = "Mki67")
VlnPlot(screp.aa,group.by = "Clone_troubleshoot",features = "Cd8b1")
VlnPlot(screp.aa,group.by = "Clone_troubleshoot",features = "Ncr1")
```
Therefore, top 2 and top 3 clones may be CD8T clones.

```{r}
DimPlot(screp.aa,group.by = "NKT_troubleshoot",order = F,alpha = 0.6,pt.size = 0.5)+scale_color_manual(values = c(pal[4:6],"grey95"))
```






```{r}
screp.aa@meta.data <- screp.aa@meta.data %>% mutate(
  celltype = case_when(
    CTaa == "CALSTPYSNNRLTL_CASSLGTGEVFF" & (celltype != "Proliferating T") ~ "CD8T", # NKT2
    CTaa == "CAVTNSNNRIFF_CAWSLSWGGSAETLYF" & (celltype != "Proliferating T") ~ "CD8T", # NKT 3
    CTaa == "CAVRPSNTNKVVF_CASSPNTEVFF" & (celltype != "Proliferating T") ~ "CD4T", # Mix
    #CTaa == "CAVNNYAQGLTF_CASSPPTGEDTQYF" & (celltype != "Proliferating T") ~ "CD8T", # NKT4
    #CTaa == "CAASERADYANKMIF_CASSRPGQGAQAPLF" & (celltype != "Proliferating T") ~ "CD8T", #NKT 5
    .default = celltype
  )
)
```

## re-do plots related to cell type labels:

```{r}
top_clone_rank <- screp.aa@meta.data %>% filter((!is.na(CTaa)) & (cloneSize == "Large (0.01 < X <= 0.1)" | cloneSize == "Medium (0.001 < X <= 0.01)")) %>% group_by(orig.ident,celltype,CTaa,clonalProportion,cloneSize) %>% summarise() %>% ungroup() %>% group_by(orig.ident) %>% slice_max(order_by = clonalProportion, n = 10) %>% ungroup()
```

```{r}
g <- ggplot(top_clone_rank) + 
  geom_bar(aes(x = reorder(CTaa, -clonalProportion), y = clonalProportion, fill = orig.ident),stat = "identity",position = "dodge")+
  facet_grid(~celltype, scale = "free")+theme_classic()+
  xlab("Clonotype by amino acid")+
  theme(axis.text.x = element_text(angle = 90,size = 8))
g
ggsave(paste0(outdir,"FigS10C_CD3r_chain_aa_top20_from_each_sample_by_prop.pdf"),plot = g, width = 20, height = 15)
```

```{r}
write.table(top_clone_rank, file = paste0(outdir,"/Top_10_clone_aa_by_sample.tsv"),sep = "\t",col.names = T,row.names = F)
```


### Gini Index


```{r}
library(DescTools)
```


```{r}
Gini_df  <- screp.aa@meta.data %>% filter((!is.na(cloneSize))) %>% group_by(orig.ident,celltype,CTaa,clonalFrequency) %>% summarise() %>% ungroup() %>% group_by(orig.ident,celltype) %>% mutate(Gini_index = Gini(clonalFrequency)) %>% ungroup()
```

```{r}
g <- ggplot(Gini_df) +
  geom_bar(aes(x = orig.ident, y = Gini_index, fill = orig.ident),stat = "identity",position = "dodge")+
  facet_wrap(~celltype,scales = "free")+xlab("Condition")+theme_classic()+ylab("Gini Index")
g
ggsave(filename = paste0(outdir,"/Fig4_Gini_index_plot.pdf"),width = 10,height = 6)
g <- ggplot(Gini_df) +
  geom_bar(aes(x = orig.ident, y = Gini_index, fill = orig.ident),stat = "identity",position = "dodge")+xlab("Condition")+theme_classic()+ylab("Gini Index")
g
ggsave(filename = paste0(outdir,"/Fig4_Gini_index_plot_bulk.pdf"),width = 10,height = 6)
```

```{r}
save(screp.aa,file = paste0(RDSdir,"lymphoid_convT_V3p3_merged_scRep_aa_relabeled.Robj"))
```

```{r}
DimPlot(screp.aa,group.by = "celltype") # sanity check
```


### proliferation clone 


```{r}
top_prolif_AA <- screp.aa@meta.data %>% group_by(orig.ident, CTaa,clonalProportion,celltype) %>% summarise() %>% ungroup() %>% group_by(orig.ident) %>% slice_max(order_by = clonalProportion, n = 20,with_ties = T) %>% ungroup()%>% arrange(clonalProportion) %>% filter(celltype == "Proliferating T")
```



```{r}
top_prolif_AA$CTaa
```


```{r}
meta <- meta %>% mutate(
  prolif_clone = case_when(
    !(CTaa %in% top_prolif_AA$CTaa) ~ "Other",
    .default = CTaa
  ),
  prolif_highlight = case_when(
    !(CTaa %in% top_prolif_AA$CTaa) ~ "Other",
    .default = "Prolif"
  )
)
```

```{r}
table(meta$prolif_clone)
```


```{r}
g <- ggplot(meta)+
  geom_point(aes(x= umap_1, y = umap_2, color = prolif_clone,size = factor(prolif_highlight,levels = c("Other","Prolif")),alpha = factor(prolif_highlight,levels = c("Other","Prolif"))))+scale_size_manual(values = c("Other" = 0.5,"Prolif" = 1))+
  facet_wrap(~orig.ident)+scale_color_manual("Clonotype",values = c(pal[1:2],pal[4:7],pal[9:10],pal[13:16],pal[18:20],pal[22:26],"grey80"))+scale_alpha_manual(values = c("Other" = 0.1,"Prolif" = 1))+theme_classic() + guides(alpha = "none",size = "none",colour = guide_legend(override.aes = list(size=3)))
g

ggsave(paste0(outdir,"Fig4_prolif_clones_on_umap.pdf"),plot = g, width = 12, height = 8)
```

```{r}
library(ggsankey)
```


```{r}
df <- meta %>% arrange(-clonalProportion) %>% filter(prolif_clone != "Other") %>% select(prolif_clone, orig.ident, celltype) %>% mutate(prolif_clone = factor(prolif_clone, levels = c("	
CALRATSSGQKLVF_CASSSGTGSERLFF","CAASTNTGKLTF_CASSADWGEDTQYF","CATEPSNTNKVVF_CASSRDPGQLYF","CAVYQGGSAKLIF_CASSLDWGGSAETLYF","CALVRGSALGRLHF_CASSQRLGGDAEQFF","CAVSIYNQGKLIF_CASRRTGGNQDTQYF","CAVMDYANKMIF_CASTRTGGYAEQFF","CAVRPSNTNKVVF_CASSPNTEVFF","CAMREGLTNTNKVVF_CASSELGVNYAEQFF","CAVNNYAQGLTF_CASSPPTGEDTQYF","CAVTNSNNRIFF_CAWSLSWGGSAETLYF","CATFLNYAQGLTF_CAWIGGAEDTQYF","CAASVETGNTGKLIF_CASSPYWGSYEQYF","CAAPGGNYKPTF_CASSGDWGVNQDTQYF","CALEQRTGNYKYVF_CASRQGGEQYF","CATESSSFSKLVF_CASSAGGSNTGQLYF","CAMREGSYTNKVVF_CASSPGQGYERLFF","CAASNAGAKLTF_CTCSDRFQAPLF","CAMREGYGSSGNKLIF_CASSWDWGGLSAETLYF","CAVMSNYNVLYF_CALGNTEVFF","Other"))) %>% make_long(prolif_clone,orig.ident,celltype)
```


```{r}
g <- ggplot(df, aes(x = x
                     , next_x = next_x
                     , node = node
                     , next_node = next_node
                     , fill = factor(node)
                     )
             )+
  geom_sankey(flow.alpha = 0.5
                      , node.color = "black"
                      ,show.legend = T)+
  theme_classic()+
  theme(legend.position = "right")+
  theme(axis.title = element_blank()
                  , axis.text.y = element_blank()
                  , axis.ticks = element_blank()  
                  , panel.grid = element_blank())+
  scale_fill_manual(values = pal)
g
ggsave(paste0(outdir,"Fig4_prolif_clones_sankey_plot.pdf"),plot = g, width = 15, height = 8)
g <- ggplot(df, aes(x = x
                     , next_x = next_x
                     , node = node
                     , next_node = next_node
                     , fill = factor(node)
                     )
             )+
  geom_sankey(flow.alpha = 0.5
                      , node.color = "black"
                      ,show.legend = T)+
  theme_classic()+
  theme(legend.position = "right")+
  geom_sankey_label(aes(label = node), size = 3, color = "white")+
  theme(axis.title = element_blank()
                  , axis.text.y = element_blank()
                  , axis.ticks = element_blank()  
                  , panel.grid = element_blank())+
  scale_fill_manual(values = pal)
g
ggsave(paste0(outdir,"Fig4_prolif_clones_sankey_plot_labeled.pdf"),plot = g, width = 15, height = 8)
```

```{r}
write.table(top_prolif_AA, file = paste0(outdir,"/Top_20_prolif_clone_aa.tsv"),sep = "\t",col.names = T,row.names = F)
```


