---
title: "0-1_preliminary_processing_and_analysis"
output: html_document
date: "2024-10-06"
---

```{r}
# define packages
library(Seurat)
library(tidyverse)
library(harmony)

library(miQC)
library(SeuratWrappers) # for miQC
library(scuttle)
```

```{r}
library(RColorBrewer)
pal <- c(brewer.pal(n = 12,name = "Set3"),brewer.pal(n = 12,name = "Paired"),brewer.pal(n = 8,name = "Accent"))
pal[2] <- "darkred"
pal[9] <- "grey25"
pal[23] <- "turquoise4"
pal[19] <- colors()[60]
pal[13] <- "darkblue"
pal[28] <- colors()[16]
pal[31] <- colors()[54] 
pal <- c(pal,colors()[c(85,84,83,89,45,31,41,42,26,29,10,139,107,108,120,109,119,121,143)])
```

```{r}
indir <- "align_outs/"
outdir <- "RDS/"
if (!dir.exists(outdir)) dir.create(outdir)

```

## load files

```{r}
samples <- list.files(indir)
```

```{r}
seu_list <- lapply(samples, function(x){
  seu_temp <- Read10X(paste0(indir,x,"/filtered_feature_bc_matrix/")) # a temporary object
  seu <- CreateSeuratObject(seu_temp)
  seu$orig.ident <- x
  seu
})
names(seu_list) <- samples # name the elements by the sample name
```


```{r}
colnames(seu_list[[1]]@meta.data)
table(seu_list[[1]]$orig.ident)
table(seu_list[[2]]$orig.ident)
table(seu_list[[3]]$orig.ident)
table(seu_list[[4]]$orig.ident)
```

## add additional meta and metrics

```{r}
seu_list <- lapply(seu_list, function(x){
  x[["percent.mt"]] <- PercentageFeatureSet(object = x, pattern = "^mt|^Mt")
  x$log10_nCount_RNA <- log10(x$nCount_RNA+1) # add a pseudcount offset to prevent log10(0)
  x$log10_nFeature_RNA <- log10(x$nFeature_RNA+1)
  x
})
```

```{r eval=FALSE}
saveRDS(seu_list,file = paste0(outdir,"filtered_sample_list.rds"))
```

## merge and process

```{r}
whole <- Reduce(function(x, y) merge(x, y), seu_list)
```

```{r}
# last minute addition of ribosomal gene percentage
whole$percent.ribo <- PercentageFeatureSet(whole,pattern="^Rp[ls]")
```



```{r}
VlnPlot(whole,features = c("log10_nFeature_RNA","log10_nCount_RNA"),group.by = "orig.ident",alpha = 0.1) #the sample quality are very similar
```


```{r}
ggplot(whole@meta.data,aes(x = log10_nCount_RNA, y = percent.mt))+
  facet_wrap(~orig.ident)+
  geom_point()+
  theme_classic()
```

```{r}
VlnPlot(whole,features = c("percent.mt","percent.ribo"),group.by = "orig.ident",alpha = 0.1) #the sample quality are very similar
```

```{r}
ggplot(whole@meta.data,aes(x = percent.ribo, y = percent.mt))+
  facet_wrap(~orig.ident)+
  geom_point()+
  theme_classic()
```
Since filtered alignment output should already have ambient contamination removed, I'll remove all cells with >0 mt.percent. I can add a label and compare the distribution pre- and post-filtering.

```{r}
whole@meta.data <- whole@meta.data %>% mutate(
  mitoQC = case_when(
    percent.mt > 0 ~ "remove",
    .default = "keep"
  )
)
```

```{r}
VlnPlot(whole,features = c("log10_nFeature_RNA","log10_nCount_RNA"),group.by = "orig.ident",alpha = 0.1,split.by = "mitoQC")

ggplot(whole@meta.data,aes(x = log10_nCount_RNA, y = percent.mt))+
  facet_wrap(~orig.ident+mitoQC, scales = "free")+
  geom_point()+
  theme_classic()
```

OK...I think 0 is too harsh...so I'll use `isOutlier` to decide the threshold.


## mito QC at sample level and merge again

```{r}
seu_list <- lapply(seu_list, function(x){
  x[["high_mito"]] <- isOutlier(x$percent.mt, type="higher", min.diff=0.5)
  
  x$percent.ribo <- PercentageFeatureSet(x,pattern="^Rp[ls]")
  x@meta.data <- x@meta.data %>% mutate(
  mitoQC = case_when(
    high_mito ~ "remove",
    .default = "keep"
  )
)
  x
}
)
```

```{r eval=FALSE}
saveRDS(seu_list,file = paste0(outdir,"filtered_sample_list.rds"))
```

```{r}
whole <- Reduce(function(x, y) merge(x, y), seu_list)
```



```{r}
VlnPlot(whole,features = c("log10_nFeature_RNA","log10_nCount_RNA"),group.by = "orig.ident",alpha = 0.1,split.by = "mitoQC")

ggplot(whole@meta.data,aes(x = log10_nCount_RNA, y = percent.mt))+
  facet_wrap(~orig.ident+mitoQC, scales = "free")+
  geom_point()+
  theme_classic()
```

```{r eval=FALSE}
saveRDS(whole,file = paste0(outdir,"whole.rds"))
```

## filter and process

```{r}
whole <- subset(whole, subset = mitoQC == "keep")
```

```{r}
ggplot(whole@meta.data,aes(x = percent.mt))+
  geom_histogram()+
  facet_wrap(~orig.ident,scales = "free",ncol = 1) 
ggplot(whole@meta.data,aes(x = log10_nFeature_RNA))+
  geom_histogram()+
  facet_wrap(~orig.ident,scales = "free",ncol = 1) 
ggplot(whole@meta.data,aes(x = log10_nCount_RNA))+
  geom_histogram()+
  facet_wrap(~orig.ident,scales = "free",ncol = 1) 
# we can see that there are still some cells suggested for discard remaining but the distribution for nCount and nFeature are about the same. To avoid noises from small number of outlier cells, we can perform hard trimming on nCount and nFeature as well.
```

```{r}
# parameters can be changed but doesn't matter much for most of the analysis
whole <- NormalizeData(object = whole, normalization.method = "LogNormalize", scale.factor = 10000)
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
#VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(whole))]
#VariableFeaturePlot(whole)

whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```


```{r message=FALSE}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20) # use harmony to correct for batch effect
whole <- RunUMAP(whole,dims = 1:20,reduction = "harmony") # use harmony embedding for downstream analysis
whole <- RunTSNE(whole,dims = 1:20,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   # iterate over a range of resolutions
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
whole <- JoinLayers(whole) # required for seurat V5
```


```{r}
DimPlot(whole,label = T,split.by = "orig.ident")
DimPlot(whole,label = T,split.by = "orig.ident",group.by = "RNA_snn_res.0.1")
```

```{r}
FeaturePlot(whole,features = "percent.mt",label = T)
FeaturePlot(whole,features = "percent.ribo",label = T) 
FeaturePlot(whole,features = "log10_nCount_RNA",label = T)
FeaturePlot(whole,features = "log10_nFeature_RNA",label = T)

``` 
### check for doublets

```{r}
VlnPlot(whole,features = c("log10_nFeature_RNA","log10_nCount_RNA"),group.by = "RNA_snn_res.0.1",alpha = 0.1)
VlnPlot(whole,features = c("log10_nFeature_RNA","log10_nCount_RNA"),group.by = "RNA_snn_res.1",alpha = 0.1)
```


```{r}
DotPlot(whole, features = c("Ptprc","P2ry12","Sall1","Hexb","Ifitm3","Isg15","Mki67"),group.by = "RNA_snn_res.0.1")
```

```{r eval=FALSE}
save(whole, file = paste0(outdir,"whole_V1p1.Robj"))
```

### round 1 QC based on markers

```{r}
Idents(whole) <- "RNA_snn_res.0.1"
FeaturePlot(whole,features = "log10_nFeature_RNA",label = T)
FeaturePlot(whole,features = "Mki67",label = T)
FeaturePlot(whole,features = "P2ry12",label = T)
FeaturePlot(whole,features = "Cd74",label = T)
FeaturePlot(whole,features = "Isg15",label = T)
FeaturePlot(whole,features = "Ifitm3",label = T)
FeaturePlot(whole,features = "Cd74",label = T)
FeaturePlot(whole,features = "Trem2",label = T)
FeaturePlot(whole,features = "Sparc",label = T)
FeaturePlot(whole,features = "Ccr6",label = T) # Dab2 , Ms4a7, Mertk, Cxcr3, Mrc1, Xcl1, Tgfb1
FeaturePlot(whole,features = "Dab2",label = T)
FeaturePlot(whole,features = "Ms4a7",label = T)
FeaturePlot(whole,features = "Mertk",label = T)
FeaturePlot(whole,features = "Cxcr3",label = T)
FeaturePlot(whole,features = "Mrc1",label = T)
FeaturePlot(whole,features = "Tgfb1",label = T) 
FeaturePlot(whole,features = "mt-Nd4l",label = T)
```

```{r}
FeaturePlot(whole,features = "Dgat1",label = T)
FeaturePlot(whole,features = "Dgat2",label = T)
FeaturePlot(whole,features = "Spp1",label = T)
FeaturePlot(whole,features = "Dkk2",label = T)
FeaturePlot(whole,features = "Gpnmb",label = T)
```


```{r}
DotPlot(whole, features = c("Ptprc","Itgae","Itgam","Itgax","Csf1","P2ry12","Sall1","Hexb","Sparc","Dab2","Ccr6","Cd74","Ifitm3","Isg15","Ms4a7","Cd9","Gpnmb","Spp1","Axl","Mertk","Cx3cr1","Tmem119","Mrc1","Xcl1","Tgfb1","Stat1","Hsp90aa1","Hspd1","Mki67","Top2a","mt-Nd4l"),group.by = "RNA_snn_res.0.1")+RotatedAxis()
```

```{r}
# Hspe1, Hsp90aa1
FeaturePlot(whole,features = "Hspe1",label = T)
FeaturePlot(whole,features = "Hsp90aa1",label = T,order = T)
FeaturePlot(whole,features = "Hspd1",label = T,order = T) #HSP60
```

```{r}
VlnPlot(whole,features = "Hsp90aa1")
VlnPlot(whole,features = "Itgam")
VlnPlot(whole,features = "Hspd1")
```

```{r}
Idents(whole) <- "RNA_snn_res.0.1"
markers <- FindAllMarkers(whole)
```
```{r eval=FALSE}
write.table(markers, file = "markers_whole_V1p1_res01.tsv",sep = "\t",col.names = T,row.names = F)
```

```{r}
top_markers <- markers %>% group_by(cluster) %>% filter(!grepl("^Gm|Rik$",gene)) %>% slice_max(order_by = avg_log2FC,n = 20)
```

```{r}
cluster2_markers <- markers %>% filter(cluster == "2")
```


```{r}
DoHeatmap(whole,features = top_markers$gene, group.by = "RNA_snn_res.0.1")
DoHeatmap(whole,features = c("Ptprc","Itgae","Itgam","Itgax","Csf1","P2ry12","Sall1","Hexb","Sparc","Dab2","Ccr6","Cd74","Ifitm3","Isg15","Ms4a7","Cd9","Gpnmb","Spp1","Axl","Mertk","Cx3cr1","Tmem119","Mrc1","Xcl1","Tgfb1","Stat1","Hsp90aa1","Hspd1","Mki67","Top2a","mt-Nd4l"), group.by = "RNA_snn_res.0.1")
```

Try removing c2


```{r}
whole <- subset(whole, idents = c(0:1,3:4))
```

```{r}
DimPlot(whole)
```


### process after filtering

```{r}
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
#VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(whole))]
#VariableFeaturePlot(whole)

whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```

```{r}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20) # use harmony to correct for batch effect
whole <- RunUMAP(whole,dims = 1:20,reduction = "harmony") # use harmony embedding for downstream analysis
whole <- RunTSNE(whole,dims = 1:20,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   # iterate over a range of resolutions
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
whole <- JoinLayers(whole) # required for seurat V5
```

```{r eval=FALSE}
save(whole, file = paste0(outdir,"whole_V1p2.Robj"))
```

```{r}
DimPlot(whole)
```

```{r}
Idents(whole) <- "RNA_snn_res.0.5"
FeaturePlot(whole,features = "log10_nFeature_RNA",label = T)
FeaturePlot(whole,features = "Mki67",label = T)
FeaturePlot(whole,features = "P2ry12",label = T)
FeaturePlot(whole,features = "Cd74",label = T)
FeaturePlot(whole,features = "Isg15",label = T)
FeaturePlot(whole,features = "Ifitm3",label = T)
FeaturePlot(whole,features = "Cd74",label = T)
FeaturePlot(whole,features = "Trem2",label = T)
FeaturePlot(whole,features = "Sparc",label = T)
FeaturePlot(whole,features = "Ccr6",label = T) # Dab2 , Ms4a7, Mertk, Cxcr3, Mrc1, Xcl1, Tgfb1
FeaturePlot(whole,features = "Dab2",label = T)
FeaturePlot(whole,features = "Ms4a7",label = T)
FeaturePlot(whole,features = "Mertk",label = T)
FeaturePlot(whole,features = "Cx3cr1",label = T)
FeaturePlot(whole,features = "Mrc1",label = T)
FeaturePlot(whole,features = "Tgfb1",label = T) 
FeaturePlot(whole,features = "mt-Nd4l",label = T)
FeaturePlot(whole,features = "Dgat1",label = T)
FeaturePlot(whole,features = "Dgat2",label = T)
FeaturePlot(whole,features = "Spp1",label = T)
FeaturePlot(whole,features = "Dkk2",label = T)
FeaturePlot(whole,features = "Gpnmb",label = T)
```
```{r}
FeaturePlot(whole,features = "percent.mt",label = T)
```


```{r}
VlnPlot(whole,features = "Dgat1",group.by  = "orig.ident")
VlnPlot(whole,features = "Dgat2",group.by = "orig.ident")
```

```{r}
DotPlot(whole, features = c("Ptprc","Itgae","Itgam","Itgax","Csf1","P2ry12","Sall1","Hexb","Sparc","Dab2","Ccr6","Cd74","Ifitm3","Isg15","Ms4a7","Cd9","Gpnmb","Spp1","Axl","Mertk","Cx3cr1","Tmem119","Mrc1","Xcl1","Tgfb1","Stat1","Hsp90aa1","Hspd1","Mki67","Top2a","mt-Nd4l"),group.by = "RNA_snn_res.0.5")+RotatedAxis()
```


