---
title: "0-1_additional_QC"
output: html_document
date: "2024-08-13"
---

```{r}
library(Seurat)
library(tidyverse)
library(miQC)
library(SeuratWrappers) # for miQC
library(harmony) # for integration
```

```{r}
# template palette for > 40 labels
library(RColorBrewer)
pal <- c(brewer.pal(n = 12,name = "Set3"),brewer.pal(n = 12,name = "Paired"),brewer.pal(n = 8,name = "Accent"))
pal[2] <- "darkred"
pal[9] <- "grey25"
pal[23] <- "turquoise4"
pal[19] <- colors()[60]
pal[13] <- "darkblue"
pal[28] <- colors()[16]
pal[31] <- colors()[54] 
pal <- c(pal,colors()[c(85,84,83,89,45,31,41,42,26,29,10,139,107,108,120,109,119,121,143)])
```

```{r}
indir <- "raw/"
outdir <- "RDS/"
if (!dir.exists(outdir)) dir.create(outdir)
```

# load object and continue QC: doublet removal

```{r}
load(file = paste0(outdir,"whole_V1.Robj"))
```

check ncounts for possible doublets

```{r}
ggplot(whole@meta.data,aes(x = log10_nCount_ADT))+
  geom_histogram()+
  facet_wrap(~miQC.keep,scales = "free",ncol = 1)
ggplot(whole@meta.data,aes(x = nCount_ADT))+
  geom_histogram()+
  facet_wrap(~miQC.keep,scales = "free",ncol = 1)
ggplot(whole@meta.data,aes(x = nCount_RNA))+
  geom_histogram()+
  facet_wrap(~miQC.keep,scales = "free",ncol = 1)
```

check distribution of the tails:
1. check tails for high nCount_RNA

```{r}
ggplot(whole@meta.data %>% filter(nCount_RNA > 25000),aes(x = nCount_ADT))+
  geom_histogram() # + facet_wrap(~orig.ident,scales = "free",ncol = 1)
ggplot(whole@meta.data %>% filter(nCount_RNA > 25000),aes(x = nCount_RNA))+
  geom_histogram() #+ facet_wrap(~orig.ident,scales = "free",ncol = 1)
```

2. check the tails for high nCount_ADT

```{r}
ggplot(whole@meta.data %>% filter(nCount_ADT > 7500),aes(x = nCount_ADT))+
  geom_histogram() # +facet_wrap(~orig.ident,scales = "free",ncol = 1)
ggplot(whole@meta.data %>% filter(nCount_ADT > 7500),aes(x = nCount_RNA))+
  geom_histogram()# + facet_wrap(~orig.ident,scales = "free",ncol = 1)
```
check number of cells exceeding some threshold and make plots on the percentage of cells

```{r}
whole@meta.data <- whole@meta.data %>% mutate(
  nCount_RNA_level = case_when(
    nCount_RNA > 25000 ~ ">25000",
    .default = "<=25000"
  ),
  nCount_ADT_level = case_when(
    nCount_ADT > 7500 ~ ">7500",
    .default = "<=7500"
))
```

```{r}
table(whole$nCount_RNA_level)
table(whole$nCount_ADT_level)
```

```{r}
table(whole@meta.data[,c("nCount_RNA_level","nCount_ADT_level")])
```
So there are 20 double-high and 65471 double-low.

```{r}
whole@meta.data <- whole@meta.data %>% mutate(
  doublet_guess = case_when(
    nCount_ADT > 7500 | nCount_RNA > 25000 ~ "doublet-like", # use OR such that only both low are considered non-doub-like
    .default = "normal-like"
  )
)
```

```{r}
qc_prop <- whole@meta.data %>% group_by(orig.ident,doublet_guess) %>% 
                            summarise(n_cat = n()) %>% 
                            ungroup() %>% 
                            group_by(orig.ident) %>%
                            mutate(prop_by_sample = n_cat*100/sum(n_cat))
```

```{r}
ggplot(qc_prop)+
  geom_bar(aes(x = orig.ident, y = prop_by_sample, fill = doublet_guess),stat = "identity",position = "stack")
```
Since the proportion of the cells seems fine, I'll use `nCount_ADT > 7500 | nCount_RNA > 25000` as the threshold and store as whole_V1p1.

```{r}
whole <- subset(whole, subset = doublet_guess == "normal-like")
```

```{r eval = FALSE}
save(whole,file = paste0(outdir,"whole_V1p1.Robj"))
```

## process V1p1 whole object and check ADT again

```{r}
load(paste0(outdir,"whole_V1p1.Robj"))
```

```{r}
# No need to normalize again since normalization is cell-specific; merging may remove the data layer, so normalization may be required after merging

# whole <- NormalizeData(object = whole, normalization.method = "LogNormalize", scale.factor = 10000)
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(whole))]
VariableFeaturePlot(whole)

whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA", "percent.mt"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
```

```{r}
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20) # use harmony to correct for batch effect
whole <- RunUMAP(whole,dims = 1:30,reduction = "harmony") # use harmony embedding for downstream analysis
whole <- RunTSNE(whole,dims = 1:30,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:30,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   # iterate over a range of resolutions
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
whole <- JoinLayers(whole)
```

```{r}
DimPlot(whole, reduction = "umap",group.by = "orig.ident",shuffle = T)
DimPlot(whole, reduction = "umap",group.by = "RNA_snn_res.0.1") # now the umap looks a little bit better
```


process ADT

```{r}
DefaultAssay(whole) <- "ADT"
whole <- NormalizeData(whole, normalization.method = "CLR", margin = 1) # normalize across features
whole <- JoinLayers(whole)
DefaultAssay(whole) <- "RNA" # switch back to RNA 
```
```{r}
DefaultAssay(whole) <- "ADT"
p1 <- FeaturePlot(whole, "CD3", cols = c("lightgrey", "red2"),order = T,slot = "data") + ggtitle("CD3 protein")
DefaultAssay(whole) <- "RNA"
p2 <- FeaturePlot(whole, "Cd3e") + ggtitle("CD3 RNA")

p1 | p2
```

```{r}
DefaultAssay(whole) <- "ADT"
p1 <- FeaturePlot(whole, "NK1.1", cols = c("lightgrey", "red2"),order = T) + ggtitle("NK1.1 protein")
DefaultAssay(whole) <- "RNA"
p2 <- FeaturePlot(whole, "Ncr1") + ggtitle("NK1.1 RNA")

p1 | p2
```
```{r}
ADT_mat <- as.data.frame(t(whole@assays$ADT@layers$data))
colnames(ADT_mat) <- rownames(whole[["ADT"]])
rownames(ADT_mat) <- colnames(whole[["ADT"]])
```

```{r}
ggplot(ADT_mat) +
  geom_histogram(aes(x = CD3))
ggplot(ADT_mat) +
  geom_histogram(aes(x = NK1.1))
ggplot(ADT_mat) +
  geom_histogram(aes(x = CCR7))
```
```{r}
ggplot(whole@meta.data,aes(x = log10_nCount_ADT))+
  geom_histogram()
ggplot(whole@meta.data,aes(x = nCount_ADT))+
  geom_histogram()
ggplot(whole@meta.data,aes(x = nCount_RNA))+
  geom_histogram()
```

Maybe I need more trimming :)))))))) The distribution has to be somewhat bimodal for ADT to work

```{r}
whole@meta.data <- whole@meta.data %>% mutate(
  nCount_RNA_level = case_when(
    nCount_RNA > 25000 ~ ">25000",
    .default = "<=25000"
  ),
  nCount_ADT_level = case_when(
    nCount_ADT > 4000 ~ ">4000",
    .default = "<=4000"
  ),
doublet_guess = case_when(
    nCount_ADT > 4000 | nCount_RNA > 25000 ~ "doublet-like", # use OR such that only both low are considered non-doub-like
    .default = "normal-like"
  ))
```

```{r}
table(whole@meta.data[,c("nCount_RNA_level","nCount_ADT_level")])
```

So I'll remove additional 470 cells

```{r}
qc_prop <- whole@meta.data %>% group_by(orig.ident,doublet_guess) %>% 
                            summarise(n_cat = n()) %>% 
                            ungroup() %>% 
                            group_by(orig.ident) %>%
                            mutate(prop_by_sample = n_cat*100/sum(n_cat))
```

```{r}
ggplot(qc_prop)+
  geom_bar(aes(x = orig.ident, y = prop_by_sample, fill = doublet_guess),stat = "identity",position = "stack")
```

Note that FE4.02 will have many cells removed

```{r}
table(whole$orig.ident)
```


```{r}
whole <- subset(whole, subset = doublet_guess == "normal-like")
```

```{r}
save(whole,file = paste0(outdir,"whole_V1p2.Robj"))
```

Ideally, you can perform sample-wise outlier detection to decide which cells to remove.


## process V1p2

```{r}
whole <- FindVariableFeatures(object = whole, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(whole) <-  VariableFeatures(whole) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(whole))]
VariableFeaturePlot(whole)

whole <- ScaleData(object = whole, features = VariableFeatures(object = whole), vars.to.regress = c("nCount_RNA", "percent.mt"))
whole <- RunPCA(object = whole,
                features =  VariableFeatures(object = whole),
                dims = 1:50)
gc()
ElbowPlot(whole,ndims = 50)
whole <- RunHarmony(object = whole, group.by.vars = c("orig.ident"), max.iter.harmony = 20) # use harmony to correct for batch effect
whole <- RunUMAP(whole,dims = 1:30,reduction = "harmony") # use harmony embedding for downstream analysis
whole <- RunTSNE(whole,dims = 1:30,reduction = "harmony")

whole <- FindNeighbors(whole, dims = 1:30,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   # iterate over a range of resolutions
  whole <- FindClusters(object = whole, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
whole <- JoinLayers(whole)
```
```{r}
DimPlot(whole, reduction = "umap",group.by = "orig.ident",shuffle = T)
DimPlot(whole, reduction = "umap",group.by = "RNA_snn_res.0.1") # now the umap looks a little bit better
```
```{r}
# check for possible immune cell types
# erythrocytes and platelets are rare but maybe also detected
DotPlot(whole,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Gata3","Rorc","Tcf7","Tox","Cd19","Flt3","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Mpo","Ccr2","Fcer1a","Fcer2a","Hbb-bs","Alas2","Pf4"),group.by = "RNA_snn_res.0.1",cluster.idents = T)+RotatedAxis()
# some of the markers are cell-state markers because cells can cluster by cell-states.
# This is important for ruling out doublet/multiplet clusters
```

```{r}
DefaultAssay(whole) <- "ADT"
whole <- NormalizeData(whole, normalization.method = "CLR", margin = 1) # normalize across features

whole <- JoinLayers(whole)
DefaultAssay(whole) <- "RNA" # switch back to RNA 
```

```{r}
DefaultAssay(whole) <- "ADT"
p1 <- FeaturePlot(whole, "CD3", cols = c("lightgrey", "red2"),order = T,slot = "data") + ggtitle("CD3 protein")
DefaultAssay(whole) <- "RNA"
p2 <- FeaturePlot(whole, "Cd3e") + ggtitle("CD3 RNA")

p1 | p2
```



```{r}
ADT_mat <- as.data.frame(t(whole@assays$ADT@layers$data))
colnames(ADT_mat) <- rownames(whole[["ADT"]])
rownames(ADT_mat) <- colnames(whole[["ADT"]])
```

```{r}
ggplot(ADT_mat) +
  geom_histogram(aes(x = CD3))
ggplot(ADT_mat) +
  geom_histogram(aes(x = NK1.1))
ggplot(ADT_mat) +
  geom_histogram(aes(x = CCR7))
```
```{r}
ggplot(whole@meta.data,aes(x = log10_nCount_ADT))+
  geom_histogram()
ggplot(whole@meta.data,aes(x = nCount_ADT))+
  geom_histogram()
ggplot(whole@meta.data,aes(x = nCount_RNA))+
  geom_histogram()
```
check cluster levels of ADT 

```{r}
ADT_agg <- AggregateExpression(whole, assays = "ADT",group.by = "RNA_snn_res.0.5",return.seurat = T)
```

```{r}
ADT_mat <- as.data.frame(log10(as.matrix(ADT_agg@assays$ADT@layers$counts)))
rownames(ADT_mat) <- rownames(ADT_agg)
colnames(ADT_mat) <- colnames(ADT_agg)
ADT_mat <- apply(ADT_mat,1,scale)
```


```{r}
RNA_agg <- AggregateExpression(whole, assays = "RNA",group.by = "RNA_snn_res.0.5",return.seurat = T)
```

```{r}
DoHeatmap(RNA_agg,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Gata3","Rorc","Tcf7","Tox","Cd19","Flt3","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Mpo","Ccr2","Fcer1a","Fcer2a","Hbb-bs","Alas2","Pf4"),draw.lines = F)
```
```{r}
DoHeatmap(ADT_agg, features = rownames(ADT_agg), draw.lines = F, slot = "data")
```
Then this is not a problem of outliers but something else.
