---
title: "0-2_secondary_processing_and_analysis"
output: html_document
date: "2024-08-27"
---

```{r}
library(Seurat)
library(tidyverse)
library(miQC)
library(SeuratWrappers) # for miQC
library(harmony) # for integration
```

```{r}
# template palette for > 40 labels
library(RColorBrewer)
pal <- c(brewer.pal(n = 12,name = "Set3"),brewer.pal(n = 12,name = "Paired"),brewer.pal(n = 8,name = "Accent"))
pal[2] <- "darkred"
pal[9] <- "grey25"
pal[23] <- "turquoise4"
pal[19] <- colors()[60]
pal[13] <- "darkblue"
pal[28] <- colors()[16]
pal[31] <- colors()[54] 
pal <- c(pal,colors()[c(85,84,83,89,45,31,41,42,26,29,10,139,107,108,120,109,119,121,143)])
```

# load data and check markers

```{r}
indir <- "RDS/"
outdir <- "R_outs/" # for storing plots if necessary
if (!dir.exists(outdir)) dir.create(outdir)
```

```{r}
load(paste0(indir,"whole_V1p2.Robj"))
```


```{r}
DimPlot(whole,group.by = "RNA_snn_res.0.1",label = T)
```

Marker panel 1:
Hbb-bs + Alas: rule out erythrocytes
Pf4: rule out platelets

The clusters aren't ideally clean yet, but we can start with lowest resolution.
c11 (Mki67+ Cd3e+ Cd8a+ Cd4-low): proliferating T cells; mostly CD8 T cells
c10: cDC1 (Flt3+ Xcr1+)
c8: pDC (Flt3+ Siglech+ Cd4+ Cd8a+)

c13 + c2: B cells (Cd19+ Cd79a+; Cd79a works better for brain B cells, but c13 is IgE receptor negative and weak in Ptprc along with weak expression in other markers, so c13 could be a multiplet cluster.)

CD4 and CD8 T cells can't be perfectly separated at the preliminary stage, so they will be temporarily grouped until the T cell subset is isolated. Some people are fine with them mingled together and the annotation can be determined based which one of them is the dominent population within the cluster.


c0,1,3,4,5,6,9: T cells (Cd3e+ Cd4+ Cd8a+);
c0: CD8 T
c1: CD4 T
c3: NKT mixture (Klrb1c+)
c4: exhausted or activated T (mainly CD8)
c5: maybe gdT (Rorc+) or CD4 Th (Th17?)
c6: Tcf7 high T cells (maybe naive CD4 and CD8)
c9: NKT mixture (Klrb1c+)

c7: ILC3 (Gata3+ Cd3e+)
c12: Rorc+; maybe another ILC


```{r}
Idents(whole) <- "RNA_snn_res.0.1"
DotPlot(whole,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Gata3","Rorc","Klrb1c","Tcf7","Tox","Cd19","Cd79a","Flt3","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Mpo","Ccr2","Fcer1a","Fcer2a","Hbb-bs","Alas2","Pf4"),group.by = "RNA_snn_res.0.1",cluster.idents = T)+RotatedAxis()
# some of the markers are cell-state markers because cells can cluster by cell-states.
# This is important for ruling out doublet/multiplet clusters
```

Decision: create a T cell subset;

```{r}
Idents(whole) <- "RNA_snn_res.0.1"
T_cells <- subset(whole, idents = c(0,1,3,4,5,6,9,7,12,11))
```

# process T cells

I temporarily decided not be worry about ribosomal RNA. If later they are suggested to affect biologically meaningful clustering for the T cells, I'll regress them out.

```{r}
T_cells <- FindVariableFeatures(object = T_cells, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(T_cells) <-  VariableFeatures(T_cells) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(T_cells))]
VariableFeaturePlot(T_cells)

T_cells <- ScaleData(object = T_cells, features = VariableFeatures(object = T_cells), vars.to.regress = c("nCount_RNA", "percent.mt"))
T_cells <- RunPCA(object = T_cells,
                features =  VariableFeatures(object = T_cells),
                dims = 1:50)
gc()
ElbowPlot(T_cells,ndims = 50)

T_cells <- RunHarmony(object = T_cells, group.by.vars = c("orig.ident"), max.iter.harmony = 20) # use harmony to correct for batch effect
T_cells <- RunUMAP(T_cells,dims = 1:30,reduction = "harmony") # use harmony embedding for downstream analysis
T_cells <- RunTSNE(T_cells,dims = 1:30,reduction = "harmony")

T_cells <- FindNeighbors(T_cells, dims = 1:30,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   # iterate over a range of resolutions
  T_cells <- FindClusters(object = T_cells, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
T_cells <- JoinLayers(T_cells)
```

```{r eval=FALSE}
save(T_cells,file = paste0(indir,"T_cells_V1p2.Robj"))
```

```{r}
DimPlot(T_cells,group.by = "RNA_snn_res.0.5",label = T)
```

```{r}
Idents(T_cells) <- "RNA_snn_res.0.5"
FeaturePlot(T_cells,features = c("Klrb1c"),label = T)
FeaturePlot(T_cells,features = c("Cd3e"),label = T)
FeaturePlot(T_cells,features = c("Cd8a"),label = T)
FeaturePlot(T_cells,features = c("Cd4"),label = T)
FeaturePlot(T_cells,features = c("Lef1","Tcf7"),label = T)
FeaturePlot(T_cells,features = c("Tox","Pdcd1"),label = T)
FeaturePlot(T_cells,features = c("Rorc"),label = T)
FeaturePlot(T_cells,features = c("Gata3"),label = T)
FeaturePlot(T_cells,features = c("Mki67"),label = T)
```

c15 is the new proliferating cluster; part of it is Cd3e-
c9 is ILC3 (Cd3e- Gata3+)
Klrb1c + (NK/NKT-like) clusters are 5,14,11

panel 1: sanity check

```{r}
Idents(T_cells) <- "RNA_snn_res.0.5"
DotPlot(T_cells,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Gata3","Rorc","Klrb1c","Tcf7","Tox","Cd19","Cd79a","Flt3","Cd44","Ncam1","Xcr1","Ncr1","Siglech","S100a8","Mpo","Ccr2","Fcer1a","Fcer2a","Hbb-bs","Alas2","Pf4"),group.by = "RNA_snn_res.0.5",cluster.idents = T)+RotatedAxis()
```

c18 is a small B cell/multiplet cluster


panel 2: more into subtypes

```{r}
Idents(T_cells) <- "RNA_snn_res.0.5"
DotPlot(T_cells,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Icos","Gata3","Rorc","Klrb1c","Lef1","Tcf7","Il2ra","Tigit","Pdcd1","Tox","Cd19","Cd79a","Cd44","Ncr1","Ccr2"),group.by = "RNA_snn_res.0.5",cluster.idents = T)+RotatedAxis()
```
c6,11,5,14 are Klrb1c+; but need to remove multiplets first.

```{r}
Idents(T_cells) <- "RNA_snn_res.0.5"
T_cells <- subset(T_cells, idents = c(0:17)) # version V1p3
```

## process V1p3


```{r}
T_cells <- FindVariableFeatures(object = T_cells, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(T_cells) <-  VariableFeatures(T_cells) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-a", VariableFeatures(T_cells))]
VariableFeaturePlot(T_cells)

T_cells <- ScaleData(object = T_cells, features = VariableFeatures(object = T_cells), vars.to.regress = c("nCount_RNA", "percent.mt"))
T_cells <- RunPCA(object = T_cells,
                features =  VariableFeatures(object = T_cells),
                dims = 1:50)
gc()
ElbowPlot(T_cells,ndims = 50)

T_cells <- RunHarmony(object = T_cells, group.by.vars = c("orig.ident"), max.iter.harmony = 20) # use harmony to correct for batch effect
T_cells <- RunUMAP(T_cells,dims = 1:30,reduction = "harmony") # use harmony embedding for downstream analysis
T_cells <- RunTSNE(T_cells,dims = 1:30,reduction = "harmony")

T_cells <- FindNeighbors(T_cells, dims = 1:30,reduction = "harmony")
for(res in seq(0.1, 1, 0.1))  {   # iterate over a range of resolutions
  T_cells <- FindClusters(object = T_cells, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
T_cells <- JoinLayers(T_cells)
```

```{r eval=FALSE}
save(T_cells,file = paste0(indir,"T_cells_V1p3.Robj"))
```

```{r}
DimPlot(T_cells,group.by = "RNA_snn_res.0.1",label = T)
```

```{r}
Idents(T_cells) <- "RNA_snn_res.0.5"
DotPlot(T_cells,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Icos","Gata3","Rorc","Klrb1c","Ncr1","Tyrobp","Lef1","Tcf7","Il2ra","Ifitm3","Tigit","Pdcd1","Tox","Cd19","Cd79a","Cd44","Ccr2"),group.by = "RNA_snn_res.0.5",cluster.idents = T)+RotatedAxis()
```

```{r}
DefaultAssay(T_cells) <- "ADT"
FeaturePlot(T_cells, "NK1.1", cols = c("lightgrey", "red2"),order = T) + ggtitle("NK1.1 protein")
```

```{r}
FeaturePlot(T_cells,features = c("percent.mt"))
FeaturePlot(T_cells,features = c("log10_nCount_RNA"))
FeaturePlot(T_cells,features = c("log10_nFeature_RNA"))
```

```{r}
prop_table <- T_cells@meta.data %>% group_by(orig.ident,RNA_snn_res.0.5) %>% summarise(cluster_size= n()) %>% ungroup() %>% group_by(orig.ident) %>% mutate(proportion = 100*cluster_size/sum(cluster_size))
```

```{r}
g <- ggplot(prop_table)+
  geom_bar(aes(x = orig.ident, y = proportion, fill = RNA_snn_res.0.5), stat = "identity",position = "stack")+scale_fill_manual(values = pal)+theme_classic()
g
ggsave(paste0(outdir,"T_cells_V1p3_res05_proportion_stack.pdf"),width = 10, height = 6)
```

```{r}
g <- ggplot(prop_table)+
  geom_bar(aes(x = orig.ident, y = proportion, fill = RNA_snn_res.0.5), stat = "identity",position = "dodge")+scale_fill_manual(values = pal)+facet_wrap(~RNA_snn_res.0.5,scales = "free")+theme_classic()+theme(axis.text.x = element_text(angle = 90))
g
ggsave(paste0(outdir,"T_cells_V1p3_res05_proportion_dodge.pdf"),width = 15,height = 10)
```
```{r}
Idents(T_cells) <- "RNA_snn_res.0.1"
DefaultAssay(T_cells) <- "RNA"
DotPlot(T_cells,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Icos","Gata3","Rorc","Klrb1c","Ncr1","Tyrobp","Lef1","Tcf7","Il2ra","Ifitm3","Tigit","Pdcd1","Tox","Cd19","Cd79a","Cd44","Ccr2"),group.by = "RNA_snn_res.0.1",cluster.idents = T)+RotatedAxis()
```

```{r}
Idents(T_cells) <- "RNA_snn_res.0.1"
p1 <- DotPlot(T_cells,features = c("Mki67","Ptprc","Cd3e","Cd8a","Cd4","Foxp3","Icos","Gata3","Rorc","Klrb1c","Ncr1","Tyrobp","Lef1","Tcf7","Il2ra","Ifitm3","Tigit","Pdcd1","Tox","Cd19","Cd79a","Cd44","Ccr2"),group.by = "RNA_snn_res.0.1",cluster.idents = T)+RotatedAxis()
p2 <- FeaturePlot(T_cells,features = c("Klrb1c"),label = T)
p3 <- FeaturePlot(T_cells,features = c("Cd3e"),label = T)
p4 <- FeaturePlot(T_cells,features = c("Cd8a"),label = T)
g <- gridExtra::grid.arrange(grobs = list(p1,p2,p3,p4),layout_matrix = rbind(c(1,1,1),c(2,3,4)))
ggsave(paste0(outdir,"T_cells_V1p3_NKT_markers.pdf"),plot = g, width = 15, height = 8)

```


```{r}
prop_table <- T_cells@meta.data %>% group_by(orig.ident,RNA_snn_res.0.1) %>% summarise(cluster_size= n()) %>% ungroup() %>% group_by(orig.ident) %>% mutate(proportion = 100*cluster_size/sum(cluster_size))

g <- ggplot(prop_table)+
  geom_bar(aes(x = orig.ident, y = proportion, fill = RNA_snn_res.0.1), stat = "identity",position = "stack")+scale_fill_manual(values = pal)+theme_classic()
g
ggsave(paste0(outdir,"T_cells_V1p3_res01_proportion_stack.pdf"),width = 10, height = 6)



g <- ggplot(prop_table)+
  geom_bar(aes(x = orig.ident, y = proportion, fill = RNA_snn_res.0.1), stat = "identity",position = "dodge")+scale_fill_manual(values = pal)+facet_wrap(~RNA_snn_res.0.1,scales = "free")+theme_classic()+theme(axis.text.x = element_text(angle = 90))
g
ggsave(paste0(outdir,"T_cells_V1p3_res01_proportion_dodge.pdf"),width = 15,height = 10)
```
```{r}
write.table(prop_table,paste0(outdir,"T_cells_V1p3_res01_proportion_table.tsv"),sep = "\t",col.names = T,row.names = F)
```



