---
title: "Check Seurat Objects for SCN"
output: html_document
date: "2024-04-08"
---

```{r}
library(Seurat)
library(harmony)
library(tidyverse)
```

```{r}
indir <- "R_outs/"
outdir <- "R_outs/"
```

# check lymphoid

```{r}
load(paste0(outdir,"lymphoid_V2p1_merged.Robj"))
```

```{r}
colnames(lymphoid@meta.data)
```
remove barcode and pANN

```{r}
lymphoid$barcodes <- NULL
lymphoid$pANN_0.25_0.09_122 <- NULL
lymphoid$pANN_0.25_0.09_163 <- NULL
lymphoid$pANN_0.25_0.09_328 <- NULL
lymphoid$pANN_0.25_0.09_384 <- NULL
lymphoid$miQC.probability <- NULL
lymphoid$NKT_TCR <- NULL
lymphoid$CD8_TCR <- NULL
lymphoid$CD4_TCR <- NULL
```
```{r}
colnames(lymphoid@meta.data)
```

```{r}
DimPlot(lymphoid,group.by = "clono_avail")
DimPlot(lymphoid,group.by = "annotation_prelim")
DimPlot(lymphoid,group.by = "celltype")
```
```{r}
save(lymphoid,file = paste0(outdir,"lymphoid_V2p1_merged.Robj"))
```

## check scRep merged lymphoid

```{r}
load(paste0(outdir,"lymphoid_V2p1_merged_scRep.Robj"))
```

```{r}
colnames(screp_lymphoid@meta.data)
```
```{r}
screp_lymphoid$barcodes <- NULL
screp_lymphoid$pANN_0.25_0.09_122 <- NULL
screp_lymphoid$pANN_0.25_0.09_163 <- NULL
screp_lymphoid$pANN_0.25_0.09_328 <- NULL
screp_lymphoid$pANN_0.25_0.09_384 <- NULL
screp_lymphoid$miQC.probability <- NULL
screp_lymphoid$NKT_TCR <- NULL
screp_lymphoid$CD8_TCR <- NULL
screp_lymphoid$CD4_TCR <- NULL
```

```{r}
colnames(screp_lymphoid@meta.data)
```
```{r}
screp_lymphoid$barcode <- NULL
```


```{r}
DimPlot(screp_lymphoid,group.by = "clono_avail")
DimPlot(screp_lymphoid,group.by = "annotation_prelim")
DimPlot(screp_lymphoid,group.by = "celltype")
DimPlot(screp_lymphoid,group.by = "cloneType",order = T)
```
```{r}
table(screp_lymphoid$cloneType)
```
```{r}
to_replace <- setNames(as.list(rep("NA",length(22:27))),colnames(screp_lymphoid@meta.data)[22:27])
to_replace[["Frequency"]] <- 0
screp_lymphoid$cloneType <- as.character(screp_lymphoid$cloneType)
screp_lymphoid@meta.data <- screp_lymphoid@meta.data %>% 
  replace_na(to_replace)
```

```{r}
screp_lymphoid$orig.ident <- as.factor(screp_lymphoid$orig.ident)
screp_lymphoid$clono_avail <- as.factor(screp_lymphoid$clono_avail)
screp_lymphoid$annotation_prelim <- as.factor(screp_lymphoid$annotation_prelim)
screp_lymphoid$celltype <- as.factor(screp_lymphoid$celltype)
screp_lymphoid$cloneType <- as.factor(screp_lymphoid$cloneType)
```

```{r}
save(screp_lymphoid,file = paste0(outdir,"lymphoid_V2p1_merged_scRep.Robj"))
```


```{r}
write.table(screp_lymphoid@meta.data,file = paste0("R_outs/","lymphoid_plotData.tsv"),sep = "\t",col.names = T,row.names = F,quote = F)
```


# conventional T cells

```{r}
load(paste0(outdir,"lymphoid_convT_V2p1_merged_scRep.Robj"))
```

```{r}
colnames(screp@meta.data)
```
```{r}
table(screp$celltype)
```

```{r}
DimPlot(screp,group.by = "annotation_prelim")
DimPlot(screp,group.by = "clono_avail")
DimPlot(screp,group.by = "cloneType")
```
```{r}
screp$barcode <- NULL
screp$barcodes <- NULL
screp$miQC.probability <- NULL
```

```{r}
to_replace <- setNames(as.list(rep("NA",length(25:31))),colnames(screp@meta.data)[25:31])
to_replace[["Frequency"]] <- 0
screp$cloneType <- as.character(screp$cloneType)
screp@meta.data <- screp@meta.data %>% 
  replace_na(to_replace)
```

```{r}
head(screp@meta.data)
```

```{r}
table(screp$clono_avail)
```
```{r}
screp$orig.ident <- as.factor(screp$orig.ident)
screp$clono_avail <- as.factor(screp$clono_avail)
screp$NKT_TCR <- as.factor(screp$NKT_TCR)
screp$CD8_TCR <- as.factor(screp$CD8_TCR)
screp$CD4_TCR <- as.factor(screp$CD4_TCR)
screp$annotation_prelim <- as.factor(screp$annotation_prelim)
screp$celltype <- as.factor(screp$celltype)
screp$cloneType <- as.factor(screp$cloneType)
screp$highlight <- as.factor(screp$highlight)
```

```{r}
save(screp,file = paste0(outdir,"lymphoid_convT_V2p1_merged_scRep.Robj"))
```


## isolate CD4

```{r}
CD4 <- subset(screp,subset = celltype == "CD4T")
```

```{r}
CD4 <- NormalizeData(object = CD4, normalization.method = "LogNormalize", scale.factor = 10000)
CD4 <- FindVariableFeatures(object = CD4, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD4) <-  VariableFeatures(CD4) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-A", VariableFeatures(CD4))]
VariableFeaturePlot(CD4)
CD4 <- ScaleData(object = CD4, features = VariableFeatures(object = CD4), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD4 <- RunPCA(object = CD4,
                features =  VariableFeatures(object = CD4),
                dims = 1:50)
gc()
ElbowPlot(CD4,ndims = 50)
```


```{r}
CD4 <- RunHarmony(object = CD4, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD4 <- RunUMAP(CD4,dims = 1:20,reduction = "harmony")
CD4 <- RunTSNE(CD4,dims = 1:20,reduction = "harmony")

CD4 <- FindNeighbors(CD4, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1.0, 0.1))  {   
  CD4 <- FindClusters(object = CD4, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```
```{r}
table(CD4$cloneType)
```
```{r}
table(CD4$celltype)
```

```{r}
DimPlot(CD4,group.by = "annotation_prelim")
DimPlot(CD4,group.by = "cloneType")
```

```{r}
CD4 <- JoinLayers(CD4)
save(CD4,file = paste0(outdir,"lymphoid_conv_CD4_V2p1_merged_scRep.Robj"))
```


## isolate CD8

```{r}
CD8 <- subset(screp,subset = celltype == "CD8T")
```

```{r}
DimPlot(CD8,group.by = "annotation_prelim")
DimPlot(CD8,group.by = "cloneType")
```
```{r}
CD8@assays$RNA@layers$scale.data.1.1.2.1 <- NULL
```


```{r}
CD8 <- NormalizeData(object = CD8, normalization.method = "LogNormalize", scale.factor = 10000)
CD8 <- FindVariableFeatures(object = CD8, selection.method = 'mean.var.plot', mean.cutoff = c(0.0125, 5), dispersion.cutoff = c(0.5, Inf))
VariableFeatures(CD8) <-  VariableFeatures(CD8) [!grepl("^Tra|^Trb|^Igh|^Igk|Mamu-A", VariableFeatures(CD8))]
VariableFeaturePlot(CD8)
CD8 <- ScaleData(object = CD8, features = VariableFeatures(object = CD8), vars.to.regress = c("nCount_RNA", "percent.mt"))
CD8 <- RunPCA(object = CD8,
                features =  VariableFeatures(object = CD8),
                dims = 1:50)
gc()
ElbowPlot(CD8,ndims = 50)
```

```{r}
CD8 <- RunHarmony(object = CD8, group.by.vars = c("orig.ident"), max.iter.harmony = 20)
CD8 <- RunUMAP(CD8,dims = 1:20,reduction = "harmony")
CD8 <- RunTSNE(CD8,dims = 1:20,reduction = "harmony")

CD8 <- FindNeighbors(CD8, dims = 1:20,reduction = "harmony")
for(res in seq(0.1, 1.0, 0.1))  {   
  CD8 <- FindClusters(object = CD8, resolution = res, print.output = 0, save.SNN = TRUE)
} 
gc()
```

```{r}
DimPlot(CD8,group.by = "annotation_prelim")
DimPlot(CD8,group.by = "RNA_snn_res.0.5")
DimPlot(CD8,group.by = "cloneType")
```

```{r}
CD8 <- JoinLayers(CD8)
save(CD8,file = paste0(outdir,"lymphoid_convT_CD8_V2p1_scRep.Robj"))
```

## check raw CD8

```{r}
load(paste0(outdir,"lymphoid_convT_CD8_V2p1.Robj"))
```

```{r}
colnames(CD8@meta.data)
```

```{r}
DimPlot(CD8,group.by = "annotation_prelim")
DimPlot(CD8,group.by = "RNA_snn_res.0.5")
```
## change strategy:

load pre-merging CD4 and CD8 subsets, integrate scRep data results into subset directly

(transferred to scRepertoire Rmd)

```{r}
load(paste0(outdir,"lymphoid_conv_CD8_V2p1_merged_scRep.Robj"))
load(paste0(outdir,"lymphoid_conv_CD4_V2p1_merged_scRep.Robj"))
```

```{r}
DimPlot(CD8,group.by = "annotation_prelim")
DimPlot(CD4,group.by = "annotation_prelim")
```
```{r}
colnames(CD8@meta.data)
colnames(CD4@meta.data)
```
```{r}
CD8$barcode <- NULL
CD8$barcodes <- NULL
CD8$miQC.probability <- NULL

CD4$barcode <- NULL
CD4$barcodes <- NULL
CD4$miQC.probability <- NULL
```

```{r}
colnames(CD8@meta.data)
```
```{r}
to_replace <- setNames(as.list(rep("NA",length(25:30))),colnames(CD8@meta.data)[25:30])
to_replace[["Frequency"]] <- 0
CD8$cloneType <- as.character(CD8$cloneType)
CD8@meta.data <- CD8@meta.data %>% 
  replace_na(to_replace) 
```


```{r}
CD8$orig.ident <- as.factor(CD8$orig.ident)
CD8$clono_avail <- as.factor(CD8$clono_avail)
CD8$annotation_prelim <- as.factor(CD8$annotation_prelim)
CD8$celltype <- as.factor(CD8$celltype)
CD8$cloneType <- as.factor(CD8$cloneType)
```

```{r}
save(CD8,file = paste0(outdir,"lymphoid_conv_CD8_V2p1_merged_scRep.Robj"))
```

```{r}
CD4$NKT_TCR <- NULL
CD4$CD4_TCR <- NULL
CD4$CD8_TCR <- NULL
colnames(CD4@meta.data)
```

```{r}
to_replace <- setNames(as.list(rep("NA",length(22:27))),colnames(CD4@meta.data)[22:27])
to_replace[["Frequency"]] <- 0
CD4$cloneType <- as.character(CD4$cloneType)
CD4@meta.data <- CD4@meta.data %>% 
  replace_na(to_replace) 
```

```{r}
CD4$orig.ident <- as.factor(CD4$orig.ident)
CD4$clono_avail <- as.factor(CD4$clono_avail)
CD4$annotation_prelim <- as.factor(CD4$annotation_prelim)
CD4$celltype <- as.factor(CD4$celltype)
CD4$cloneType <- as.factor(CD4$cloneType)
```

```{r}
save(CD4,file = paste0(outdir,"lymphoid_conv_CD4_V2p1_merged_scRep.Robj"))
```

# Second Round

Still had problem with lymphoid; previous there were problems with whole datasets.

## lymphoid
```{r}
load(paste0(outdir,"lymphoid_V2p1_merged_scRep.Robj"))
```

```{r}
DimPlot(screp_lymphoid,group.by = "cloneType")
DimPlot(screp_lymphoid,group.by = "celltype")
```

```{r}
table(screp_lymphoid$cloneType)
```
```{r}
colnames(screp_lymphoid@meta.data)
```
```{r}
table(screp_lymphoid$celltype)
```


```{r}
screp_lymphoid$clono_avail <- NULL
```


```{r}
meta <- screp_lymphoid@meta.data
UMAP <- screp_lymphoid@reductions$umap@cell.embeddings
TSNE <- screp_lymphoid@reductions$tsne@cell.embeddings
meta = cbind(UMAP,TSNE,meta)
```

```{r}
ggplot(meta,aes(x = umap_1,y=umap_2,color = cloneType))+
  geom_point()
```

```{r}
write.table(meta,file = paste0(outdir,"lymphoid_plotData.tsv"),sep = "\t",col.names = T,row.names = F,quote = F)
```

```{r}
save(screp_lymphoid,file = paste0(outdir,"lymphoid_V2p1_merged_scRep.Robj"))
```


```{r}
plotData <- read.table(file = paste0(outdir,"lymphoid_plotData.tsv"), sep = "\t",header = T)
```


```{r}
ggplot(meta,aes(x = umap_1,y=umap_2,color = cloneType))+
  geom_point()
```

```{r}
NK <- subset(screp_lymphoid,subset = celltype == "NK")
```

```{r}
table(NK$cloneType)
```
## conventional T

```{r}
load(paste0(outdir,"lymphoid_convT_V2p1_merged_scRep.Robj"))
```

```{r}
meta <- screp@meta.data
UMAP <- screp@reductions$umap@cell.embeddings
TSNE <- screp@reductions$tsne@cell.embeddings
meta = cbind(UMAP,TSNE,meta)
```

```{r}
ggplot(meta,aes(x = umap_1,y=umap_2,color = cloneType))+
  geom_point()
```

```{r}
write.table(meta,file = paste0(outdir,"conventional_T_plotData.tsv"),sep = "\t",col.names = T,row.names = F,quote = F)
```

## CD4

```{r}
load(paste0(outdir,"lymphoid_conv_CD4_V2p1_merged_scRep.Robj"))
```

```{r}
meta <- CD4@meta.data
UMAP <- CD4@reductions$umap@cell.embeddings
TSNE <- CD4@reductions$tsne@cell.embeddings
meta = cbind(UMAP,TSNE,meta)
```

```{r}
ggplot(meta,aes(x = umap_1,y=umap_2,color = cloneType))+
  geom_point()
```

```{r}
write.table(meta,file = paste0(outdir,"CD4_T_plotData.tsv"),sep = "\t",col.names = T,row.names = F,quote = F)
```


## CD8

```{r}
load(paste0(outdir,"lymphoid_conv_CD8_V2p1_merged_scRep.Robj"))
```

```{r}
meta <- CD8@meta.data
UMAP <- CD8@reductions$umap@cell.embeddings
TSNE <- CD8@reductions$tsne@cell.embeddings
meta = cbind(UMAP,TSNE,meta)
```

```{r}
ggplot(meta,aes(x = umap_1,y=umap_2,color = cloneType))+
  geom_point()
```

```{r}
write.table(meta,file = paste0(outdir,"CD8_T_plotData.tsv"),sep = "\t",col.names = T,row.names = F,quote = F)
```

# check PC

```{r}
load(paste0(indir,"lymphoid_V2p1_merged_scRep.Robj"))
```

```{r}
print(screp_lymphoid[["pca"]], dims = 1:5)
```

```{r}
VizDimLoadings(screp_lymphoid, dims = 1:2, reduction = "pca",nfeatures = 40,balanced = T)
```

```{r}
TopFeatures(object = screp_lymphoid[["pca"]], dim = 1, nfeatures = 80, balanced = TRUE)
```


```{r}
g <- DotPlot(screp_lymphoid,features = TopFeatures(object = screp_lymphoid[["pca"]], dim = 1, nfeatures = 80, balanced = TRUE), group.by = "RNA_snn_res.0.6",cluster.idents = T) + RotatedAxis()
ggsave("PC1_top80_balanced_lymphoid_res06_DotPlot.pdf",plot = g,width = 20,height = 10)
```

```{r}
g <- DotPlot(screp_lymphoid,features = TopFeatures(object = screp_lymphoid[["pca"]], dim = 1, nfeatures = 80, balanced = TRUE), group.by = "RNA_snn_res.0.5",cluster.idents = T) + RotatedAxis()
ggsave("PC1_top80_balanced_lymphoid_res05_DotPlot.pdf",plot = g,width = 20,height = 10)
```

## perform DE for 

```{r}
colnames(screp_lymphoid@meta.data)
```

```{r}
DimPlot(screp_lymphoid,group.by = "celltype")
DimPlot(screp_lymphoid,group.by = "annotation_prelim")
DimPlot(screp_lymphoid,group.by = "RNA_snn_res.0.6",label = T)
```
```{r}
# 10, 1 , 6
# 16, 3, 0, 15, 2
# Actually to make things simplier, just run DE and GSEA over res.0.6
Idents(screp_lymphoid) <- "RNA_snn_res.0.6"
DE_genes <- FindAllMarkers(screp_lymphoid)
```


```{r}
library(msigdbr)
library(fgsea)
library(org.Mm.eg.db)
library(Seurat) # youb know why
library(tidyverse) # base package for data organization and plotting
library(RColorBrewer) # for making palettes
library(gridExtra) # for arranging multiple plots
library(data.table) # well...
```

```{r prepare gene sets}
H_df<- msigdbr(species = "Mus musculus", category = "H")
Imm_df <- msigdbr(species = "Mus musculus", category = "C7",subcategory = "IMMUNESIGDB")
Celltype_df <- msigdbr(species = "Mus musculus", category = "C8")
C2_KEGG <- msigdbr(species = "Mus musculus", category = "C2",subcategory = "CP:KEGG")
C2_React <- msigdbr(species = "Mus musculus", category = "C2",subcategory = "CP:REACTOME")


fgsea_sets_Hallmark <- H_df %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_Imm <- Imm_df %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_KEGG <- C2_KEGG %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_REACT <- C2_React %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_Celltype <- Celltype_df %>% split(x = .$gene_symbol, f = .$gs_name)

```

```{r}
fgsea_res <- function(cell_type_vector, DE_genes, choose_n = 100, choose_tops = F,gene_set = fgsea_sets_Hallmark,mode = "combined",outdir){
  if(!dir.exists(outdir)) {dir.create(outdir)}
  fgsea_res <- list()
  #DE_selected <- DE_genes %>% group_by(cluster) %>% filter(p_val_adj < 0.05 & abs(avg_log2FC) > 1 & pct.1 > 0.1) %>% mutate(metric = sign(avg_log2FC)*(-log10(p_val_adj))) %>% filter(!is.infinite(metric))
  DE_selected <- DE_genes %>% mutate(metric = sign(avg_log2FC)*(-log10(p_val_adj))) %>% filter(!is.infinite(metric))
  if (mode == "simple"){
    DE_selected <- DE_genes %>% mutate(metric = avg_log2FC) %>% filter(!is.infinite(metric))
  }
  for (i in cell_type_vector) {
    DE <- DE_selected[DE_selected$cluster == i,] %>% arrange(desc(metric)) %>% dplyr::select(gene,metric)
    if (choose_tops){
      DE_top <- DE %>% group_by(cluster) %>% slice_max(order_by = metric, n = choose_n)
      DE_tail <- DE %>% group_by(cluster) %>% slice_min(order_by = metric, n = choose_n)
      DE <- rbind(DE_top,DE_tail)
    }
    DE <- DE %>% arrange(desc(metric)) %>% dplyr::select(gene,metric)
    rankings <- DE$metric
    names(rankings) <- DE$gene
    #print(rankings)
    plot(rankings)
    title(main = i)
    fgseaRes <- fgsea(gene_set, stats = rankings,nPermSimple = 10000)
    fgseaResTidy <- fgseaRes %>%
      as_tibble() %>% filter(padj < 0.05 ) %>%
      arrange(desc(NES))
  fgsea_res[[i]] <- fgseaResTidy
  fwrite(fgseaResTidy, file=paste0(outdir,i,"_fgsea.tsv"), sep="\t", sep2=c("", " ", ""))
  } # need to add name of gene sets to the file name
   return(fgsea_res)
}
```

```{r}
outdir <- "functional_analysis/"
if (!dir.exists(outdir)){
  dir.create(outdir)
}
```

```{r}
fgsea_hallmark <- fgsea_res(as.character(unique(DE_genes$cluster)),DE_genes,gene_set = fgsea_sets_Hallmark,mode = "simple",outdir = paste0(outdir,"Hallmark/"))

Hallmark_vector <- names(fgsea_sets_Hallmark)
celltype_vector <- as.character(unique(DE_genes$cluster))
table_res <- data.frame(pathway = Hallmark_vector)
for (i in celltype_vector){
  table_tmp <- fgsea_hallmark[[i]]
  table_celltype <- table_tmp[table_tmp$pathway %in% Hallmark_vector,c("pathway","NES")]
  colnames(table_celltype) <- c("pathway",i)
  table_res <- full_join(table_res,table_celltype,by = "pathway")
}

table_res[is.na(table_res)] = 0
table_res <- table_res[apply(table_res[,2:(length(unique(DE_genes$cluster))+1)],1,var) != 0,]

write.table(table_res,file = paste0(outdir,"Hallmark/","fgsea_res06_Hallmark_consolidated_filtered.tsv") ,sep = "\t",row.names = F,quote = F)
```

```{r}
fgsea_REACT <- fgsea_res(as.character(unique(DE_genes$cluster)),DE_genes,gene_set = fgsea_sets_REACT,mode = "simple",outdir = paste0(outdir,"Reactome/"))

REACT_vector <- names(fgsea_sets_REACT)
celltype_vector <- as.character(unique(DE_genes$cluster))
table_res <- data.frame(pathway = REACT_vector)
for (i in celltype_vector){
  table_tmp <- fgsea_REACT[[i]]
  table_celltype <- table_tmp[table_tmp$pathway %in% REACT_vector,c("pathway","NES")]
  colnames(table_celltype) <- c("pathway",i)
  table_res <- full_join(table_res,table_celltype,by = "pathway")
}

table_res[is.na(table_res)] = 0
table_res <- table_res[apply(table_res[,2:(length(unique(DE_genes$cluster))+1)],1,var) != 0,]

write.table(table_res,file = paste0(outdir,"Reactome/","fgsea_res06_Reactome_consolidated_filtered.tsv") ,sep = "\t",row.names = F,quote = F)
```

```{r}
fgsea_celltype <- fgsea_res(as.character(unique(DE_genes$cluster)),DE_genes,gene_set = fgsea_sets_Celltype,mode = "simple",outdir = paste0(outdir,"Celltype/"))

CellType_vector <- names(fgsea_sets_Celltype)
celltype_vector <- as.character(unique(DE_genes$cluster))
table_res <- data.frame(pathway = CellType_vector)
for (i in celltype_vector){
  table_tmp <- fgsea_celltype[[i]]
  table_celltype <- table_tmp[table_tmp$pathway %in% CellType_vector,c("pathway","NES")]
  colnames(table_celltype) <- c("pathway",i)
  table_res <- full_join(table_res,table_celltype,by = "pathway")
}

table_res[is.na(table_res)] = 0
table_res <- table_res[apply(table_res[,2:(length(unique(DE_genes$cluster))+1)],1,var) != 0,]

write.table(table_res,file = paste0(outdir,"Celltype/","fgsea_res06_Celltype_consolidated_filtered.tsv") ,sep = "\t",row.names = F,quote = F)
```

```{r}
fgsea_imm <- fgsea_res(as.character(unique(DE_genes$cluster)),DE_genes,gene_set = fgsea_sets_Imm,mode = "simple",outdir = paste0(outdir,"Immunogenic/"))

imm_vector <- names(fgsea_sets_Imm)
celltype_vector <- as.character(unique(DE_genes$cluster))
table_res <- data.frame(pathway = imm_vector)
for (i in celltype_vector){
  table_tmp <- fgsea_imm[[i]]
  table_celltype <- table_tmp[table_tmp$pathway %in% imm_vector,c("pathway","NES")]
  colnames(table_celltype) <- c("pathway",i)
  table_res <- full_join(table_res,table_celltype,by = "pathway")
}

table_res[is.na(table_res)] = 0
table_res <- table_res[apply(table_res[,2:(length(unique(DE_genes$cluster))+1)],1,var) != 0,]

write.table(table_res,file = paste0(outdir,"Immunogenic/","fgsea_res06_Immunogenic_consolidated_filtered.tsv") ,sep = "\t",row.names = F,quote = F)
```

```{r}
fgsea_KEGG <- fgsea_res(as.character(unique(DE_genes$cluster)),DE_genes,gene_set = fgsea_sets_KEGG,mode = "simple",outdir = paste0(outdir,"KEGG/"))

KEGG_vector <- names(fgsea_sets_KEGG)
celltype_vector <- as.character(unique(DE_genes$cluster))
table_res <- data.frame(pathway = KEGG_vector)
for (i in celltype_vector){
  table_tmp <- fgsea_KEGG[[i]]
  table_celltype <- table_tmp[table_tmp$pathway %in% KEGG_vector,c("pathway","NES")]
  colnames(table_celltype) <- c("pathway",i)
  table_res <- full_join(table_res,table_celltype,by = "pathway")
}

table_res[is.na(table_res)] = 0
table_res <- table_res[apply(table_res[,2:(length(unique(DE_genes$cluster))+1)],1,var) != 0,]

write.table(table_res,file = paste0(outdir,"KEGG/","fgsea_res06_KEGG_consolidated_filtered.tsv") ,sep = "\t",row.names = F,quote = F)
```

