---
title: "Within-cell type DE"
output: html_document
---

```{r}
library(Seurat)
library(tidyverse)
library(data.table)
library(MAST)
```

```{r}
indir = "./R_outs/"
outdir = "./R_outs/"
```

```{r}
load(paste0(outdir,"whole_V2p4.Robj"))
```

```{r}
table(whole$celltype)
```

```{r}
table(whole$hashtag)
table(paste(whole$orig.ident,whole$hashtag,sep = "_"))
length(paste(whole$orig.ident,whole$hashtag,sep = "_"))
```

```{r}
whole$condition <- paste(whole$orig.ident,whole$hashtag,sep = "_")
```

```{r}
whole@meta.data <- whole@meta.data %>%
  mutate(annotation_V2 = case_match(annotation,
    "Mono/mac_1: M1-like 1" ~ "Mono/Mac_1",
    "Mono/mac_2: Mertk- Ccr2+ M1-like" ~ "Mono/Mac_1",
    "Mono/mac_3: Mertk+ Ccr2- Arg- M2-like" ~ "Mono/Mac_2",
    "Mono/Mac_4: Arg1+ Ifitm1+" ~ "Mono/Mac_3",
    "Mono/Mac_5: Mertk+ Ccr2- Arg1+ M2-like" ~ "Mono/Mac_4",
    "Mono/Mac_6: Rsad2 high M1-like" ~ "Mono/Mac_5",
    "Mono/Mac_7: M1-like 2" ~ "Mono/Mac_1",
    "Mono/Mac_8: Mertk - Ccr2+ Mrc1- Il1r2- M2-like" ~ "Mono/Mac_1",
    "Mono/Mac_9: Mertk+ Ccr2+ M2-like" ~ "Mono/Mac_1",
    .default = annotation))
```

```{r}
table(whole@meta.data$annotation_V2)
```


```{r}
save(whole, file=paste0(outdir,"whole_V2p4.Robj"))
```

### perform DE among CAF

```{r}
CAF <- subset(whole,subset = celltype == "CAF")
```

```{r}
table(CAF$annotation)
```


```{r}
Idents(CAF) <- CAF$annotation

CAF_markers <- FindAllMarkers(CAF)


```

```{r}
write.table(CAF_markers,
              paste0("R_outs/","CAF_annotation_markers.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```


### DE among DC

```{r}
DC <- subset(whole,subset = celltype == "DC")
table(DC$annotation)
```

```{r}
Idents(DC) <- DC$annotation

markers <- FindAllMarkers(DC)

write.table(markers,
              paste0("R_outs/","DC_annotation_markers.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```

### DE Neutrophils

```{r}
Neutrophil <- subset(whole,subset = celltype == "Neutrophil")
table(Neutrophil$annotation)
```

```{r}
Idents(Neutrophil) <- Neutrophil$annotation

markers <- FindAllMarkers(Neutrophil)

write.table(markers,
              paste0("R_outs/","Neutrophil_annotation_markers.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```

### DE among mono/mac

```{r}
mono_mac <- subset(whole,subset = celltype == "Mono/Mac")

```

```{r}
table(mono_mac$annotation_V2)
```


```{r}
Idents(mono_mac) <- mono_mac$annotation_V2

markers <- FindAllMarkers(mono_mac)

write.table(markers,
              paste0("R_outs/","mono_mac_annotation_markers_V2.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```

```{r}
test <- read.table(file = paste0("R_outs/","mono_mac_annotation_markers.tsv"),header = T,sep = "\t")
```

### functional analysis of DE results

```{r}
library(presto)
library(msigdbr)
library(fgsea)
library(ReactomeGSA)
library(gridExtra)
```

#### mono/mac

```{r}
DE_table <- read.table(file = paste0("R_outs/","mono_mac_annotation_markers_V2.tsv"),header = T,sep = "\t")
```


```{r}
plot_HallmarkGSEA <- function(cluster_list,celltype,DE_table,padj = 0.01,log2FC = 1){
  # here the cluster list is the list of annotation to go over
  n <- length(cluster_list)
  plts <- list(n)
  fgsea_table <- data.frame()
  DE <- DE_table %>% filter(p_val_adj < padj & avg_log2FC >= log2FC)
  for (i in seq(1:n)){
    cluster_DE_genes<- DE %>%
    dplyr::filter(cluster == cluster_list[i]) %>%
    arrange(desc(avg_log2FC)) %>%
    dplyr::select(gene,avg_log2FC)
    ranks<- deframe(cluster_DE_genes)
    fgseaRes <- fgseaMultilevel(fgsea_sets_Hallmark, stats = ranks,scoreType = "pos",eps = 0)
    fgseaResTidy <- fgseaRes %>%
      as_tibble() %>%
      arrange(desc(NES))
    ggplot(fgseaResTidy %>% filter(padj < 0.01) %>% head(n= 25), aes(reorder(pathway, NES), NES)) +
    geom_col(aes(fill= NES < 7.5)) +
    coord_flip() +
    labs(x="Hallmark Pathway", y="Normalized Enrichment Score",
       title=paste0("",cluster_list[i])) +
    theme_minimal()
    plts[[i]] <- ggplot(fgseaResTidy %>% filter(padj < 0.01) %>% head(n= 25), aes(reorder(pathway, NES), NES)) +
    geom_col(aes(fill= NES < 7.5)) +
    coord_flip() +
    labs(x="Hallmark Pathway", y="Normalized Enrichment Score",
       title=paste0("",cluster_list[i])) +
    theme_minimal()
    fgsea_table_tmp <- cbind(fgseaResTidy,cluster = rep(cluster_list[i],nrow(fgseaResTidy)))
    fgsea_table <- rbind(fgsea_table,fgsea_table_tmp)
  }
  result <- list(plts = plts,table = fgsea_table)
  return(result)
}
```

```{r}
cbind(fgseaResTidy,cluster = rep(cluster_list[1],nrow(fgseaResTidy)))
```

```{r}
H_df<- msigdbr(species = "Mus musculus", category = "H")
Imm_df <- msigdbr(species = "Mus musculus", category = "C7")

fgsea_sets_Hallmark <- H_df %>% split(x = .$gene_symbol, f = .$gs_name)
fgsea_sets_Imm <- Imm_df %>% split(x = .$gene_symbol, f = .$gs_name)
```


```{r}
# Doublecheck helper function
res <- plot_HallmarkGSEA(unique(DE_table$cluster),"mono_mac",DE_table,log2FC = 1.25)
g <- gridExtra::grid.arrange(grobs = res$plts,ncol = 3)
ggsave(filename = paste0(outdir,"Hallmark_GSEA_barplots_mono_mac.pdf"),plot = g,height = 60,width = 80,units = "cm")
```

```{r}
fgsea_table <- as.data.frame(res$table)
```


```{r}
write.table(fgsea_table,
              paste0("./R_outs/","mono_mac_annotation_fgsea.tsv"),
              sep="\t",
              quote=F,
              row.names=F)
```


```{r}
plotEnrichment(fgsea_sets_Hallmark[["HALLMARK_INTERFERON_GAMMA_RESPONSE"]],
               ranks) + labs(title="HALLMARK_INTERFERON_GAMMA_RESPONSE")
```




## Direct and indirect effect DE

```{r}
table(whole$condition)
```

We need to first define a function for DE to be repeated over different condition combination while looping over annotation. We also need to generate a table for each annotation with information on number of cells from each condition.

Note that most cells from lymphoid are reporter -,it is thus meaningless to perform this DE over lymphoid cells.

```{r}
sp <- subset(whole,subset = (hashtag == "Reporter+") | (hashtag == "Reporter-"))
```

```{r}
table(sp$condition)
```

```{r}
save(sp, file=paste0(outdir,"whole_V2p4_annotated_SP.Robj"))
```


```{r}
sp_condition_table <- sp@meta.data %>% group_by(condition,annotation) %>%  
  summarise(n = n()) %>% mutate(total_by_condition = sum(n)) %>% mutate(percentage = n*100/total_by_condition) %>% ungroup() %>% mutate(total = sum(n))
```

```{r}
g <- ggplot(sp_condition_table,aes(x = condition, y = percentage,fill = condition))+
  geom_bar(stat = "identity", position = 'dodge')+
  facet_wrap(~annotation,scales = "free")+guides(fill=guide_legend(title="Condition"))+ ylab("Percentage of cells with same condition")+
  theme(
  plot.title = element_blank(),
  axis.title.x = element_blank())+RotatedAxis()
ggsave(filename = paste0(outdir,"Proportion_barplots_condition_by_annotation.pdf"), plot = g, units = "cm",height = 30,width = 50)
```

```{r}
g <- ggplot(sp_condition_table,aes(x = condition, y = n,fill = condition))+
  geom_bar(stat = "identity", position = 'dodge')+
  facet_wrap(~annotation,scales = "free")+guides(fill=guide_legend(title="Condition"))+ ylab("Percentage of cells with same condition")+
  theme(
  plot.title = element_blank(),
  axis.title.x = element_blank())+RotatedAxis()
ggsave(filename = paste0(outdir,"number_barplots_condition_by_annotation.pdf"), plot = g, units = "cm",height = 30,width = 50)
```

```{r}
sp_condition_table <- sp@meta.data %>% group_by(annotation,condition) %>%  
  summarise(n = n()) %>% mutate(total_by_annotation = sum(n))  %>% mutate(percentage = n*100/total_by_annotation) %>% ungroup() %>% mutate(total = sum(n))
```


```{r}
write.table(sp_condition_table,file = paste0(outdir,"SP_annotation_table_by_condition.tsv"),sep = "\t",quote = F,row.names = F)
```

```{r}
g <- ggplot(sp_condition_table,aes(x = condition, y = percentage,fill = condition))+
  geom_bar(stat = "identity", position = 'dodge')+
  facet_wrap(~annotation,scales = "free")+guides(fill=guide_legend(title="Condition"))+ ylab("Percentage of cells with the same annotation")+
  theme(
  plot.title = element_blank(),
  axis.title.x = element_blank())+RotatedAxis()
ggsave(filename = paste0(outdir,"Proportion_barplots_annotation_by_condition.pdf"), plot = g, units = "cm",height = 30,width = 50)
```





```{r}
de_clusters <- function(current_cluster) {
  groups <- paste0(conditions,"_",current_cluster)
  both_markers <- FindMarkers(lymphoid, ident.1=groups[2], ident.2=groups[1], test.use="MAST", min.pct=0.00, logfc.threshold = 0)
  marker_genes <- row.names(both_markers)
  mutate(both_markers, gene = marker_genes, cluster = current_cluster)
  
}
```

